\PassOptionsToPackage{colorlinks=true,allcolors=black}{hyperref}
% REQ-FILE: The above line, \PassOptionsToPackage{} must be very first line.

\documentclass[11pt]{article}

% ============================================================
% Annotation visibility toggle
% Build annotated PDF by defining \ANNOTATED (e.g., via latexmk flag).
% Requires xcolor loaded first, then use \textcolor and \color commands.
% ============================================================
\newif\ifShowAnnotations
\ifdefined\ANNOTATED
  \ShowAnnotationstrue
\else
  \ShowAnnotationsfalse
\fi

% Encoding and font setup for cross-platform reproducibility
\usepackage[T1]{fontenc}
\usepackage{lmodern}

% Improve readability for dense theoretical material
\usepackage{setspace}
\onehalfspacing

% Mathematical notation and table formatting for formal semantics
\usepackage{amsmath, amssymb}
\usepackage{array}
\usepackage[table]{xcolor}
\usepackage{colortbl}
\usepackage{booktabs}
\usepackage{graphicx}

% Theorem environments (best-practice, widely standard)
\usepackage{amsthm}

% ------------------------------------------------------------
% Numbering convention:
% - All theorem-like statements share one counter, numbered by section.
% - Definition/Example/Remark/Note also share the same counter family.
%   (This is the most common convention in math/logic/category theory writing.)
% ------------------------------------------------------------

% Italic body text: results
\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{conjecture}[theorem]{Conjecture}

% Claim varies; most standard is to share the theorem counter
\newtheorem{claim}[theorem]{Claim}

% Upright body text: definitions and examples
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}

% Upright body text: remarks/notes
\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{note}[theorem]{Note}

% Optional: unnumbered versions for front-matter previews/callouts
% (This is also standard when you want a preview without numbering.)
\newtheorem*{theorem*}{Theorem}
\newtheorem*{lemma*}{Lemma}
\newtheorem*{proposition*}{Proposition}
\newtheorem*{corollary*}{Corollary}
\newtheorem*{conjecture*}{Conjecture}
\newtheorem*{claim*}{Claim}
\newtheorem*{definition*}{Definition}
\newtheorem*{example*}{Example}
\newtheorem*{remark*}{Remark}
\newtheorem*{note*}{Note}

% Bibliography management and hyperlink support
\usepackage{url}
\usepackage{natbib}
% load hyperref before bookmark to avoid warnings
\usepackage{hyperref}
\usepackage{bookmark}

% Support multiple author affiliations
\usepackage{authblk}

% Improve typographic quality and line breaking
\usepackage{microtype}

% Categorical and commutative diagrams
\usepackage{tikz}
\usepackage{tikz-cd}
\usetikzlibrary{arrows.meta,calc,fit,positioning,shapes.geometric}

% Callout boxes for examples and emphasis
\usepackage[most]{tcolorbox}

% Keywords macro for structured abstract metadata
\providecommand{\keywords}[1]{\textbf{Keywords: } #1}

% Notation for Raw vs Canonical categories in EP/CEE
\newcommand{\Raw}{\mathsf{Raw}}
\newcommand{\Canon}{\mathsf{Canon}}

% Reusable figure callout box for visual emphasis
\newcommand{\FigureCallout}[2]{%
  \begin{tcolorbox}[
    colback=gray!5,
    colframe=black!40,
    title={#1},
    fonttitle=\bfseries,
    arc=3pt,
    boxrule=0.5pt,
    width=\linewidth,
    enhanced,
    breakable
  ]
  #2
  \end{tcolorbox}
}
% Notation for stability-critical 
\newcommand{\stabilitycritical}{stability-critical}

% ============================================================
% Annotations (REQ / WHY / OBS) scan-friendly
% ============================================================
\newcommand{\AnnColor}{gray} % xcolor color name
\newlength{\AnnHangIndent}
\setlength{\AnnHangIndent}{1.8em} % hanging indent amount

% Single annotation line/block:
% - small gray text
% - bold label REQ(P0) then content
% - hanging indent for scanning
\newcommand{\AnnInline}[3]{%
  \ifShowAnnotations
    \par\begingroup
      \footnotesize
      \color{\AnnColor}%
      \noindent
      \hangindent=\AnnHangIndent
      \hangafter=1
      \textbf{#1(#2):}~#3\par
    \endgroup
    \vspace{0.15\baselineskip}%
  \fi
}
\newcommand{\REQ}[2]{\AnnInline{REQ}{#1}{#2}}
\newcommand{\WHY}[2]{\AnnInline{WHY}{#1}{#2}}
\newcommand{\OBS}[2]{\AnnInline{OBS}{#1}{#2}}

% Paragraph spacing prioritizes readability over traditional indentation
\setlength{\parskip}{0.75em}
\setlength{\parindent}{0em}

% Compact list formatting for dense conceptual content
\usepackage{enumitem}
\setlist[itemize]{itemsep=0.2em, topsep=0.2em, parsep=0em, partopsep=0em}

% Defensive re-definition ensures commands exist if loaded earlier
\providecommand{\Raw}{\mathrm{Raw}}
\providecommand{\Canon}{\mathrm{Canon}}

% ============================================================
\title{Substrate Stability Under Persistent Disagreement: \\
Structural Constraints for \\ Neutral Ontological Substrates }
% ============================================================

\author[1,2]{Denise M. Case}
\affil[1]{Northwest Missouri State University, Computer Science and Information Systems, Maryville, MO, USA}
\affil[2]{Civic Interconnect, Ely, MN, USA}

\date{\today}

\begin{document}

\maketitle
\vspace{-1em}

% Explicitly mark preprint status
\begin{tcolorbox}[colback=gray!5, colframe=black!20, boxrule=0.3pt]
  \textbf{Preprint Notice.}
  This preprint has not undergone peer review.
  It is shared to support community discussion, transparency research, and early technical evaluation.
\end{tcolorbox}

\begin{abstract}
Modern data systems increasingly operate under conditions of persistent legal,
political, and analytic disagreement.
In such settings, interoperability cannot rely on shared interpretation,
negotiated semantics, or centralized authority.
Instead, representations must function as neutral substrates that preserve
stable reference across incompatible extensions.

This paper investigates the structural constraints imposed on ontological
design by this requirement.
Building on a neutrality framework that treats interpretive non-commitment and
stability under extension as explicit design constraints, we ask what minimal
ontological structure is forced if accountability relationships are to remain
referable and comparable under disagreement.
Minimality here is not mere parsimony:
a reduction is admissible only if it does not 
reintroduce \stabilitycritical{} distinctions 
as hidden roles, flags, or contextual predicates.

We establish a conditional lower-bound result: any ontology capable of
supporting accountability under persistent disagreement must realize at least
six distinct identity-and-persistence regimes.
We further show that a construction with exactly six such regimes is sufficient
to satisfy the stated requirements without embedding causal or normative
commitments in the substrate.
The result is not a proposal for a universal ontology, but a constraint on what
is possible when neutrality and stable reference are treated as non-negotiable
design goals.
\end{abstract}


% =============================================================================
% Theorem Preview 
% =============================================================================

\bigskip
\noindent\textbf{Statement of Result (Preview).}
\emph{Under explicit neutrality and stability constraints, we show that attempts to
simplify ontological structure by collapsing distinctions inevitably reintroduce
those distinctions as hidden interpretive mechanisms.
When such hidden regimes are excluded, accountability under persistent
disagreement requires a minimal set of six distinct identity-and-persistence regimes:
fewer do not satisfy the stated neutrality and stability requirements, 
and more are unnecessary.}

\medskip
\noindent
Informally: Neutrality is achieved by
externalizing explanatory and evaluative structure.
To function as a substrate for disagreement, an ontology must remain
agnostic with respect to causal and normative interpretation.

\medskip
\noindent
The formal statement and proof appear in Section~\ref{sec:interpretation}.

\keywords{
  formal ontology;
  ontological neutrality;
  accountability;
  neutral substrates;
  extension stability;
  causal and normative commitment
}

\ifShowAnnotations
  \clearpage
  \input{sections/00_contract}
  \clearpage
\fi

\input{sections/01_intro}           
\input{sections/02_assumptions}        
\input{sections/03_requirements}    
\input{sections/04_necessity}  
\input{sections/05_sufficiency}   
\input{sections/06_interpretation}
\input{sections/07_related}
\input{sections/08_conclusion}     


\section*{Statements and Declarations}

\subsection*{Author Contributions}
The author was solely responsible for the conception, analysis, and writing of this manuscript.

\subsection*{Declaration of Conflicting Interest}
The author declares no potential conflicts of interest
with respect to the research, authorship,
and/or publication of this article.

\subsection*{Funding}
This research received no external funding.

\subsection*{Data Availability}
No datasets were generated or analyzed
during the current study.

\subsection*{Ethical Approval}
Not applicable.

\subsection*{Consent to Participate}
Not applicable.

\subsection*{Consent for Publication}
Not applicable.

\appendix
\clearpage

\input{sections/99_appendixG}   % Glossary

\ifShowAnnotations
  \clearpage
  \input{sections/99_appendixZ}   % Trace (internal)
\fi

\bibliographystyle{plainnat}
\bibliography{bibliography/bibliography}
\end{document}