\PassOptionsToPackage{colorlinks=true,allcolors=black}{hyperref}
% REQ-FILE: The above line, \PassOptionsToPackage{} must be very first line.

\documentclass[11pt]{article}

% Encoding and font setup for cross-platform reproducibility
\usepackage[T1]{fontenc}
\usepackage{lmodern}

% Improve readability for dense theoretical material
\usepackage{setspace}
\onehalfspacing
\usepackage{caption}
\captionsetup[table]{font=small,skip=4pt}

% Mathematical notation and table formatting for formal semantics
\usepackage{amsmath, amssymb}
\usepackage{array}
\usepackage[table]{xcolor}
\usepackage{colortbl}
\usepackage{booktabs}
\usepackage{graphicx}

% Theorem environments (best-practice, widely standard)
\usepackage{amsthm}

% Transformation classifications
\newcommand{\PRS}{\textsf{PRS}}
\newcommand{\BRK}{\textsf{BRK}}
\newcommand{\NEU}{\textsf{NEU}}
\newcommand{\NA}{\textsf{N/A}}

% Identity regimes
\newcommand{\OBL}{\mathrm{OBL}}
\newcommand{\OCC}{\mathrm{OCC}}
\newcommand{\REC}{\mathrm{REC}}

\newcommand{\LOC}{\mathrm{LOC}}
\newcommand{\OBJ}{\mathrm{OBJ}}

\newcommand{\SCOPEE}{\mathrm{SCOPE\text{-}E}}
\newcommand{\SCOPES}{\mathrm{SCOPE\text{-}S}}

\newcommand{\RULEC}{\mathrm{RULE\text{-}C}}
\newcommand{\RULES}{\mathrm{RULE\text{-}S}}

% Formal notation
\newcommand{\Rspace}{\mathcal{R}}
\newcommand{\Fset}{\mathcal{F}}
\newcommand{\Class}{\mathrm{Class}}
\newcommand{\basis}{\mathrm{basis}}
\newcommand{\Substrate}{\mathcal{S}}
\newcommand{\Frameworks}{\mathcal{F}}
\newcommand{\SubstrateRef}{\mathsf{Ref}(\Substrate)}

% ------------------------------------------------------------
% Numbering convention:
% - All theorem-like statements share one counter, numbered by section.
% - Definition/Example/Remark/Note also share the same counter family.
%   (This is the most common convention in math/logic/category theory writing.)
% ------------------------------------------------------------

% Italic body text: results
\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{conjecture}[theorem]{Conjecture}

% Claim varies; most standard is to share the theorem counter
\newtheorem{claim}[theorem]{Claim}

% Upright body text: definitions and examples
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}

% Upright body text: remarks/notes
\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{note}[theorem]{Note}
\newtheorem{convention}[theorem]{Convention}

% Optional: unnumbered versions for front-matter previews/callouts
% (This is also standard when you want a preview without numbering.)
\newtheorem*{theorem*}{Theorem}
\newtheorem*{lemma*}{Lemma}
\newtheorem*{proposition*}{Proposition}
\newtheorem*{corollary*}{Corollary}
\newtheorem*{conjecture*}{Conjecture}
\newtheorem*{claim*}{Claim}
\newtheorem*{definition*}{Definition}
\newtheorem*{example*}{Example}
\newtheorem*{remark*}{Remark}
\newtheorem*{note*}{Note}
\newtheorem*{convention*}{Convention}

% Bibliography management and hyperlink support
\usepackage{url}
\usepackage{natbib}
\bibpunct{(}{)}{;}{a}{}{ }

% load hyperref before bookmark to avoid warnings
\usepackage{hyperref}
\usepackage{bookmark}

% Support multiple author affiliations
\usepackage{authblk}

% Improve typographic quality and line breaking
\usepackage{microtype}

% Keywords macro for structured abstract metadata
\providecommand{\keywords}[1]{\textbf{Keywords: } #1}

% Paragraph spacing prioritizes readability over traditional indentation
\setlength{\parskip}{0.75em}
\setlength{\parindent}{0em}

% Compact list formatting for dense conceptual content
\usepackage{enumitem}
\setlist[itemize]{itemsep=0.2em, topsep=0.2em, parsep=0em, partopsep=0em}

% =============================================================================
% PREFERRED CITATION SPECIFICATION
%
% @techreport{case2026regimes,
%   author      = {Case, Denise M.},
%   title       = {Referential Regimes: Transformation-Invariant Identity for Neutral Substrates},
%   institution = {Northwest Missouri State University},
%   year        = {2026},
%   doi         = {10.48550/arXiv.2601.16152},
%   url         = {https://doi.org/10.48550/arXiv.2601.16152}
% }
% =============================================================================


% ============================================================
\title{Referential Regimes:  \\
Transformation-Invariant Identity \\
for Neutral Substrates}
% ============================================================
\author{Denise M. Case}
\affil{Northwest Missouri State University, Computer Science and Information Systems, Maryville, MO, USA}
\date{} % do not change

\begin{document}

\maketitle
\vspace{-1em}

\begin{abstract}
  Data systems increasingly operate under persistent
  legal, political, and analytic disagreement,
  where no single interpretive authority can be assumed.
  A \emph{neutral substrate} provides stable shared reference
  without requiring agreement about causal or normative interpretation:
  it fixes reference structurally and leaves interpretation to
  extension layers, following the neutrality-by-design constraint
  developed in \emph{Neutral Substrates}~\citep{case2026neutral}.

  This paper derives the identity structure such a substrate requires,
  using a small transformation algebra rather than a list of kinds.
  We treat identity as transformation-invariance:
  a referent is individuated by the admissible operations that preserve it.
  Once the routine operations that record systems perform are fixed,
  including re-expression, annotation, refinement, decomposition,
  reassignment, forking, and provenance extension,
  individuation becomes a structural calculation.

  That calculation exposes a mismatch between reference kinds and
  identity regimes. An accountability substrate must refer to six kinds
  of thing: obligation-bearers, rules, occurrences, scopes, records, and
  plain referents.
  Three of those kinds are assigned more than one admissible basis.
  A scope may be fixed by its extension or by its structure;
  a rule by its content or by its structure;
  a plain referent by its locus or by its object.
  A single routine operation separates each pair:
  decomposition separates the two scope readings,
  refinement the two rule readings, and forking
  the two plain-referent readings.

  Counting reference kinds gives six;
  transformation analysis forces at least nine identity regimes.
  The three additional regimes add no new kind to the inventory.
  They follow from how identity behaves under operations
  the substrate cannot avoid performing, and any attempt to collapse them
  reappears as a hidden regime that persistent disagreement
  cannot hold fixed.

  The result is a conditional lower bound, monotone in the relevant sense:
  a richer transformation basis may leave the bound unchanged
  or force further splits, but it cannot erase a distinction
  already witnessed here. Its content is diagnostic.
  The additional regimes are the distinctions upper-ontology
  implementations may leave to the modeler, and that cannot be recovered
  under persistent disagreement without hidden regimes.
\end{abstract}

\keywords{
  formal ontology;
  identity and persistence;
  neutral substrates;
  extension stability;
  accountability;
  causal commitment;
  normative commitment
}

\input{sections/01_intro}
\input{sections/02_background}
\input{sections/03_formal_setting}
\input{sections/04_reference_inventory}
\input{sections/05_algebra}
\input{sections/06_nine_regimes}
\input{sections/07_lower_bound}
\input{sections/08_determinancy}
\input{sections/09_application}
\input{sections/60_related_work}
\input{sections/70_limits}
\input{sections/80_conclusion}

% =============================================================================
\section*{Statements and Declarations}
% =============================================================================

\subsection*{Acknowledgments} The author thanks the anonymous reviewers
of an earlier version of this work for comments that helped improve
the framing, organization, and presentation of the work.

\subsection*{Author Contributions}
The author is the sole contributor to this work and
is responsible for all aspects of the research, authorship, and publication.

\subsection*{Use of AI-Assisted Tools}
AI-assisted tools were used for editing, formatting, and consistency checking.
The author reviewed all suggestions and is solely responsible for the content.

\subsection*{Declaration of Conflicting Interest}
The author declares no potential conflicts of interest with respect to the
research, authorship, and publication of this work.

\appendix
\clearpage
\input{sections/99_appendixA}   % pairwise collapse enumeration

\begin{thebibliography}{99}
  \small

  \bibitem[Arp et al.(2015)]{arp2015building}
  R.\ Arp, B.\ Smith, and A.\ D.\ Spear.
  \emph{Building Ontologies with Basic Formal Ontology}.
  MIT Press, 2015.

  \bibitem[Bowker and Star(1999)]{bowkerstar1999}
  G.\ C.\ Bowker and S.\ L.\ Star.
  \emph{Sorting Things Out: Classification and Its Consequences}.
  MIT Press, Cambridge, MA, 1999.

  \bibitem[Case(2026)]{case2026neutral}
  D.\ M.\ Case.
  \emph{Neutral Substrates: A Design Constraint for Interoperability under Persistent Disagreement}.
  arXiv preprint \href{https://arxiv.org/abs/2601.14271}{arXiv:2601.14271}, 2026.

  \bibitem[Edwards(2011)]{edwards2011}
  P.\ N.\ Edwards.
  \emph{A Vast Machine: Computer Models, Climate Data, and the Politics of Global Warming}.
  MIT Press, 2011.

  \bibitem[Gangemi et al.(2002)]{gangemi2002dolce}
  A.\ Gangemi, N.\ Guarino, C.\ Masolo, A.\ Oltramari, and L.\ Schneider.
  Sweetening ontologies with DOLCE.
  In \emph{EKAW 2002}, volume 2473 of LNCS, pages 166--181. Springer, 2002.

  \bibitem[Grieves and Vickers(2016)]{grieves2016}
  M.\ Grieves and J.\ Vickers.
  Digital twin: Mitigating unpredictable, undesirable emergent behavior in complex systems.
  In \emph{Transdisciplinary Perspectives on Complex Systems}, pages 85--113. Springer, 2016.

  \bibitem[Guarino(1998)]{guarino1998roles}
  N.\ Guarino.
  On the ontological status of roles.
  \emph{Applied Ontology}, 1998.

  \bibitem[Guarino and Welty(2002)]{guarinowelty2002}
  N.\ Guarino and C.\ A.\ Welty.
  Evaluating ontological decisions with OntoClean.
  \emph{Communications of the ACM}, 45(2):61--65, 2002.

  \bibitem[Guizzardi(2005)]{guizzardi2005}
  G.\ Guizzardi.
  \emph{Ontological Foundations for Structural Conceptual Models}.
  PhD thesis, University of Twente, 2005.

  \bibitem[Haslanger(2012)]{haslanger2012}
  S.\ Haslanger.
  \emph{Resisting Reality: Social Construction and Social Critique}.
  Oxford University Press, 2012.

  \bibitem[Hodges(1997)]{hodges1997}
  W.\ Hodges.
  \emph{A Shorter Model Theory}.
  Cambridge University Press, 1997.

  \bibitem[ISO/IEC(2018)]{isoiec24707}
  ISO/IEC.
  Information technology --- Common Logic (CL): A framework for a family of logic-based languages.
  ISO/IEC 24707:2018, 2018.

  \bibitem[Kutz et al.(2010)]{kutz2010}
  O.\ Kutz, T.\ Mossakowski, and D.\ L\"ucke.
  Carnap, Goguen, and the hyperontologies: Logical pluralism and heterogeneous structuring in ontology design.
  \emph{Logica Universalis}, 4(2):255--333, 2010.

  \bibitem[Longino(1990)]{longino1990}
  H.\ E.\ Longino.
  \emph{Science as Social Knowledge: Values and Objectivity in Scientific Inquiry}.
  Princeton University Press, 1990.

  \bibitem[Masolo et al.(2003)]{masolo2003wonderweb}
  C.\ Masolo, S.\ Borgo, A.\ Gangemi, N.\ Guarino, and A.\ Oltramari.
  WonderWeb Deliverable D18: Ontology Library.
  IST Project 2001-33052 WonderWeb, 2003.

  \bibitem[Moreau and Missier(2013)]{provdm2013}
  L.\ Moreau and P.\ Missier (eds.).
  PROV-DM: The PROV Data Model.
  W3C Recommendation, 30 April 2013.

  \bibitem[Pearl(2009)]{pearl2009}
  J.\ Pearl.
  \emph{Causality: Models, Reasoning, and Inference}.
  Cambridge University Press, 2nd edition, 2009.

  \bibitem[Smith and Ceusters(2015)]{smithceusters2015bfo}
  B.\ Smith and W.\ Ceusters.
  Basic Formal Ontology 2.0: Specification and user's guide.
  \emph{Applied Ontology}, 10(1):1--77, 2015.

  \bibitem[Wilkinson et al.(2016)]{wilkinson2016}
  M.\ D.\ Wilkinson et al.
  The FAIR Guiding Principles for scientific data management and stewardship.
  \emph{Scientific Data}, 3:160018, 2016.

\end{thebibliography}

\end{document}
