% !TeX root = se200_identity_regimes.tex
% =============================================================================
% Glossary: Identity Regimes and Structural Constraints
% =============================================================================

\section*{Appendix B. Glossary of Terms}
\addcontentsline{toc}{section}{Appendix B. Glossary of Terms}

\textit{This glossary defines key terms used in the paper.}

\begin{description}[style=nextline,leftmargin=1.5cm]

  \item[Applicability Context]
        An enduring referent that delimits where, to whom, or under what conditions
        an authority-bearing structure applies.
        Applicability contexts persist independently of specific occurrences
        and must support nesting and overlap.

  \item[Hidden Regime]
        An identity- or persistence-relevant distinction that is not represented
        as an explicit ontological kind,
        but is instead encoded via roles, flags, contextual predicates,
        or interpretive conventions supplied by extension layers.
        Hidden regimes are excluded under the stability optimization adopted here.

  \item[Identity-and-Persistence Regime]
        A structured set of identity and persistence conditions
        that determine how a class of entities can be stably referenced
        over time within an ontology.
        An identity-and-persistence regime specifies what counts as the same entity,
        what changes are compatible with continued identity, and
        how references to that entity remain stable across
        datasets, jurisdictions, or observational contexts.
        In a neutral substrate, such regimes provide the identity
        and persistence structure upon which interpretive frameworks operate.
        This paper identifies exactly six regimes that are sufficient
        to support stable reference under the neutrality constraint.

  \item[Individuation]
        The specification of the identity and persistence conditions
        that determine how entities are distinguished from one another
        and tracked as the same or different across time and contexts.
        Individuation criteria govern how entities are identified,
        how they persist through change, and how they remain referable
        across interpretive frameworks.

  \item[Interpretive Extension]
        Any explanatory, normative, analytic, or evaluative structure
        added atop the substrate that supplies meaning, causation,
        or judgment without altering substrate identity conditions.

  \item[Lower-Bound Result]
        A necessity claim establishing that no representation
        with fewer than a specified number of identity-and-persistence regimes
        can satisfy the stated requirements under the adopted optimization criterion.

  \item[Neutrality Constraint]
        The requirement that the substrate itself
        remain pre-causal and pre-normative,
        permitting incompatible interpretations
        without requiring revision of entity identity or structure.

  \item[Neutral Substrate]
        A foundational ontology that satisfies neutrality,
        enabling stable use across multiple,
        potentially incompatible interpretive frameworks.
        A neutral substrate excludes causal and normative commitments
        from the foundational layer,
        supporting pluralism and accountability.

  \item[Optimization Criterion]
        The objective function fixed for this paper:
        to minimize explicit ontological structure
        subject to neutrality, stable reference,
        and the exclusion of hidden identity regimes.
        Minimality is defined relative to stability, not syntactic brevity.

  \item[Persistent Disagreement]
        A condition in which interpretive, normative,
        or analytic incompatibilities are expected to endure over time,
        rather than converge through negotiation, evidence, or governance.

  \item[Stable Reference]
        The property that entities remain referable
        as the same entities across time and across incompatible extensions,
        without reclassification or reinterpretation of identity criteria.

  \item[Sufficiency Construction]
        An explicit ontological partition that satisfies all stated requirements
        using exactly the minimal number of identity-and-persistence regimes
        established by the lower-bound result.

  \item[Tight Bound]
        A result showing that the same number of regimes
        is both necessary and sufficient
        under the stated assumptions and optimization objective.

  \item[Time-Indexed Occurrence]
        An entity whose identity is inseparable from its temporal realization
        and provenance.
        Occurrences do not persist beyond their happening
        and are distinct from enduring entities.

\end{description}
