% !TeX root = se200_identity_regimes.tex
\section{Introduction: Structure Under Disagreement}
\label{sec:intro}

% ============================================================
% A. Problem Context and Motivation
% ============================================================

% A1. General Motivation / problem statement
Modern data systems must often support stable reference and
accountability across settings characterized by persistent legal,
political, and analytic disagreement.% TODO NEEDS CITE
% A2. Add examples to illustrate the problem concretely (A support)
Three recurring settings make the structural requirements concrete.
First, in multi-jurisdiction regulatory reporting,
a single corporate entity operates under overlapping legal regimes
that individuate obligations, violations, and responsible parties differently;
audit trails must remain comparable across jurisdictions
without presupposing which regime governs.
Second, in legislative history tracking,
statutory authority is amended, delegated, and contested over time,
requiring stable reference to normative structures
independently of the occurrences that enact or challenge them.
Third, in cross-institutional provenance analysis,
research or administrative records are produced,
transferred, and re-analyzed across institutions whose
classificatory and causal assumptions diverge,
yet must remain referable as the same records
across those divergent frameworks.
These cases are not offered as empirical evidence;
they function as structural diagnostics that make concrete
the requirements of Section~\ref{sec:requirements}.

% A3. Standard tools are not designed for this
Upper ontologies and general-purpose representational frameworks
including BFO~\citep{arp2015bfo}, DOLCE~\citep{gangemi2002dolce},
and PROV-O~\citep{lebo2013prov} are expressive and well-motivated
for their intended purposes.
They support rich domain modeling, causal attribution,
and normative classification.
However, expressiveness is not neutrality.
These frameworks were designed to represent the world accurately
within an interpretive context, not to remain agnostic across
incompatible interpretive contexts simultaneously.
When deployed as shared substrates under persistent disagreement
without additional constraints,
their embedded commitments such as what counts as an agent,
what individuates an event, and
what constitutes an authority,
can become sites of interpretive conflict
rather than stable reference points.

% A4. This is a special case
The problem addressed here is distinct from the general
ontology design problem.
The objective is not representational adequacy within a given framework
but substrate stability across frameworks that cannot be assumed to converge.
No single interpretive authority can be presupposed,
and semantic convergence cannot be enforced
without excluding legitimate participants or perspectives.
Nevertheless, shared use of data remains necessary:
entities must be referable, actions traceable, and
records comparable across frameworks that may diverge
in assumptions, methods, or evaluative standards.
Recent work confirms that systems without explicit structural commitments
produce unstable and context-dependent categorical behavior~\citep{kohler2025mercurial},
further motivating the need for substrates that enforce stable reference
structurally rather than relying on interpretive consistency.

% A5. Trying to apply standard cases
A natural response might be to reduce ontological commitment by collapsing
distinctions and representing entities, occurrences, authorities, and records
as a single undifferentiated type, differentiated only by role predicates
(e.g., \emph{isAgent}, \emph{isOccurrence}, \emph{isAuthority})
or using contextual flags supplied at query or extension time.
This achieves surface parsimony but not structural neutrality.
Such encodings relocate essential distinctions from the substrate
to interpretive mechanisms that are not represented at the substrate level,
and whose consistent application requires precisely the shared commitments
that persistent disagreement denies.
We refer to these as \emph{hidden regimes},
structural discriminations that are operationally necessary
but ontologically unacknowledged.
They are incompatible with the design objective,
because they make referential stability
contingent on interpretive consensus that cannot be assumed.

% ============================================================
% B. Summarize Design Constraints from Paper 1
% ============================================================

% B1. neutrality design constraint
The preceding analysis identifies a design objective that is distinct
from the general problem of ontology construction.
We adopt a neutrality framework developed in prior work~\citep{case2026neutrality},
treating ontological neutrality as a structural design constraint.
Neutrality is understood as interpretive non-commitment at the substrate level
combined with stability under incompatible extension:
a substrate must permit independent extensions to introduce
causal, normative, or analytic structure without requiring revision of the
underlying representation.
Explanatory and evaluative structure is externalized;
the substrate is constrained to remain pre-causal and pre-normative.

% B2. layer boundary and interpretive agnosticism
The substrate is not intended to adjudicate causal or normative claims.
Its role is limited to preserving referential comparability
across independently evolving extensions.
All causal explanation, normative evaluation, and
domain-specific interpretation are delegated to extension layers.

% ============================================================
% C. Stability (Defined Precisely)
% ============================================================

Stable reference in this work concerns regime-level persistence conditions,
not immunity to transformation or versioning at interpretive layers.
The substrate must preserve the referential integrity of entities and records
even as their interpretation evolves in incompatible ways across extensions.
Persistent disagreement, as used in this work, is defined by four conditions:
\begin{itemize}
\item Extensions may introduce incompatible causal accounts.
\item Extensions may introduce incompatible normative attributions.
\item Extensions cannot be forced into convergence.
\item Substrate revision is not permitted to resolve interpretive conflict.
\end{itemize}
Under these conditions, the substrate must remain interpretively agnostic
while preserving referential stability across extensions.
% Ship of Theseus clarification
Stability is a property of the identity-and-persistence regime,
rather than any particular token or interpretive assignment within it.

% ============================================================
% D. Minimality as Structural Constraint
% ============================================================

% optimization objective - Reviewer 1 Concern 3
Within this setting, minimality is not a reduction in the number of primitives.
Apparent simplifications may achieve surface parsimony
by reintroducing essential distinctions indirectly
through roles, flags, contextual typing, or other internal discriminators
interpreted at extension layers.
Such encodings do not eliminate structure;
they relocate it to mechanisms that presuppose shared interpretive commitments.
Under persistent disagreement, those commitments cannot be assumed.
Any reduction that preserves stability only by
deferring identity-and-persistence distinctions
to extension layers therefore fails the design objective.
Minimality is thus defined as the exclusion of hidden coordination mechanisms:
all operationally necessary identity-and-persistence distinctions
must be explicit and structurally enforced at the substrate level.

% ============================================================
% E. Target Capacity: Accountability
% ============================================================

% CRITICAL STRUCTURAL QUESTION AND Define Target Capacity
% What structure is required to support accountability under these constraints?
Accountability is the specific capacity discussed in this paper.
Accountability analysis requires that a neutral substrate
support stable reference to participants,
authority structures, regulated occurrences, applicability contexts, and
descriptive records,
without embedding causal claims or normative judgments in the substrate itself.
These are not introduced as a proposed ontological partition,
but as representational capacities whose simultaneous support
places structural constraints on any neutral substrate.
Whether they can be realized within fewer identity-and-persistence regimes
is addressed in Section~\ref{sec:necessity};
the question is whether a neutral substrate can support them at all,
and if so, with what irreducible structural commitments.
Three diagnostic cases (multi-jurisdiction regulatory reporting,
legislative history tracking, and cross-institutional provenance)
recur throughout the paper to make these requirements concrete;
they are structural diagnostics, not empirical evidence.

% ============================================================
% F. Contribution
% ============================================================

The central contribution of this paper is a conditional bound on ontological
structure given the design objectives outlined above.
Under these neutrality and stability constraints, we show that any ontology
capable of supporting accountability under persistent disagreement must
realize at least six distinct identity-and-persistence regimes.
We further show that a construction realizing exactly six such regimes is
sufficient to satisfy the stated requirements without introducing hidden or
implicit structure.
The result is not a proposal for a universal ontology but a constraint on
what is structurally possible under a specific optimization objective.

% VERY IMPORTANT Scope Containment (F refinement)
Alternative objectives prioritize different tradeoffs,
such as interpretive convergence, domain-specific efficiency,
or context-dependent classification at the substrate level;
these are not discussed here.

The contribution of this paper is a constraint result.
The question addressed is structural rather than taxonomic:
if a representational substrate must support accountability
across independently evolving interpretations
while preserving stable reference to shared entities and records,
what identity-and-persistence structure, if any, is unavoidable?

Treating neutrality and extension-stable reference as explicit design
constraints and grounding the analysis in concrete accountability workflows
(regulatory reporting, legislative history tracking, and
cross-institutional provenance),
the paper derives a conditional lower bound on the number of regimes required.

The remainder of the paper proceeds as follows.
Section~\ref{sec:assumptions} states the assumptions, neutrality constraints,
and optimization criterion.
Section~\ref{sec:requirements} formalizes the structural requirements imposed
by accountability analysis.
Section~\ref{sec:necessity} derives the lower bound.
Section~\ref{sec:sufficiency} presents a construction achieving this bound.
Section~\ref{sec:interpretation} discusses the scope and implications.
Section~\ref{sec:related} situates the work relative to existing frameworks.
Section~\ref{sec:conclusion} concludes.
