% !TeX root = se200_identity_regimes.tex
\section{Assumptions and Design Constraints}
\label{sec:assumptions}

This section fixes the problem setting for the remainder of the paper.
All results that follow are conditional on the assumptions stated here.
The purpose is not to argue that these assumptions are universally correct,
but to make explicit the constraints under which a neutral substrate for
accountability is possible.
These constraints determine how entities may be individuated within such
a substrate.

\subsection{Persistent Disagreement and Extension Without Coordination}

We assume environments in which disagreement is not an anomaly to be
resolved, but a persistent condition.
Legal interpretation, policy evaluation, and analytic methodology may remain
incompatible across communities and over time.
Accordingly, no single interpretive authority or shared semantic framework
is assumed at the substrate level.

We further assume that extensions to the substrate may occur without
coordination.
Independent actors may add explanatory, normative, or analytic structure,
and such extensions must not require revision of the substrate itself.
This assumption does not rely on
global semantic agreement, centralized governance, or
negotiated meaning as preconditions for interoperability.
Under these assumptions,
substrate-level admissibility must not depend
on interpretive agreement among extensions.

\subsection{Stability Under Persistent Disagreement}

Stable reference in this work concerns regime-level identity and persistence conditions,
not immunity to transformation or versioning at interpretive layers.
The substrate must preserve the referential integrity of entities and records
even as their interpretation evolves in incompatible ways across extensions.
Persistent disagreement, as used in this work, is defined by four conditions:
\begin{itemize}
\item Extensions may introduce incompatible causal accounts.
\item Extensions may introduce incompatible normative attributions.
\item Extensions cannot be forced into convergence.
\item Substrate revision is not permitted to resolve interpretive conflict.
\end{itemize}
Under these conditions, the substrate must remain interpretively agnostic
while preserving referential stability across extensions.
% Ship of Theseus clarification
Stability is a property of the identity-and-persistence regime that governs individuation,
rather than any particular token or interpretive assignment within it.

\subsection{Neutrality Constraint}

This work assumes substrate neutrality and stability
as a design constraint rather than as a philosophical claim.
A neutral substrate does not assert why
events occur, whether outcomes are desirable, or
how actions ought to be evaluated.
Instead, it provides stable referents and structural relations upon which
external interpretive layers may operate.

Neutrality further requires stability under incompatible extension
at the level of identity conditions and admissible relations.
If two extensions disagree in interpretation, evaluation, or explanation,
that disagreement must not force reclassification of entities, alteration
of identity conditions, or revision of the substrate.
Any representation whose admissibility conditions depend on interpretive
agreement does not satisfy this constraint.

\subsection{Stable Reference Requirement}
\label{sec:stable_reference_requirement}
For a substrate to support accountability, entities must remain referable
over time and across extensions.
Identity conditions and admissible relations must not depend
on interpretive convergence.
If an entity must change its ontological status in order to participate in new
relationships or analyses, stable reference is not preserved.
This requirement is not satisfied by representations in which identity is
determined contextually or retroactively by interpretive layers.
The domain consequences of this requirement are illustrated in
Section~\ref{sec:requirements} under R1.

This requirement excludes representations in which \stabilitycritical{}
distinctions are encoded only as roles, flags, or contextual predicates.
Such mechanisms rely on interpretive conventions supplied by extension layers
and therefore permit divergent admissibility judgments under persistent
disagreement.

The stability requirement is intentionally strong.
Engineering systems may address identity pressure through contextual
or negotiated identity schemes, and in such settings strong reference
stability may not be attainable in practice.
The present analysis instead examines the case
where such mechanisms are unavailable and
stable reference must be preserved independently of interpretive convergence.

\subsection{Optimization Criterion}

The optimization objective of this paper is to minimize explicit ontological
structure subject to the constraints above.
Minimality is not defined as
the smallest possible number of primitives in the abstract,
but as the smallest number of explicit identity-and-persistence
regimes required to preserve neutrality and stable reference.

A proposed reduction constitutes a genuine simplification only if it does not
reintroduce the collapsed distinctions through hidden discriminators,
interpretive mechanisms, or role predicates.

The minimality criterion adopted here is analytical rather than prescriptive:
it is introduced to determine what ontological structure
is unavoidable when neutrality and stability under persistent disagreement
are treated as design constraints.

Alternative optimization objectives prioritize different tradeoffs, such as
interpretive coordination, domain-specific efficiency, or context-dependent
classification at the substrate level.
The results presented here apply only to substrates optimized for
interoperability under persistent disagreement.

These assumptions and constraints determine the design space
within which a neutral accountability substrate must operate.
The following section states the representational capacities that any
substrate must satisfy under these conditions.
