% !TeX root = ../se200_referential_regimes.tex

% =============================================================================
\section{Background: Neutral Substrates}
\label{sec:background}
% =============================================================================

This paper builds on the \emph{Neutral Substrates (NS)} paper~\citep{case2026neutral},
which states the design constraint for a shared representational base that must
remain usable under persistent interpretive disagreement.
This paper analyzes one aspect of that constraint:
the identity and persistence structure needed for stable reference to survive
ordinary transformations of records.

The definitions in this section are inherited from that work.
They are reproduced here to make the dependency clear.
Local labels are used in this paper, while the definition titles indicate
their source.

% -----------------------------------------------------------------------------
\subsection{Inherited Definitions}
\label{subsec:ns-inherited-definitions}
% -----------------------------------------------------------------------------

\begin{definition}[Substrate, from NS]
  \label{def:ns-substrate}
  A \emph{substrate} $\Substrate$ is a shared representational base providing
  stable reference for entities, occurrences, and institutional artifacts
  across a class of admissible interpretive frameworks $\Frameworks$.
  Stable reference consists of individuation, co-reference, and persistence.
\end{definition}

A substrate is not the whole record system.
It is the foundational layer on which further interpretive claims may be layered,
including causal or normative interpretations,
evidentiary assessments, explanatory accounts, and institutional determinations.
Its purpose is to make shared reference possible without forcing agreement about
every interpretation of the referent.
A substrate may identify an event, entity, decision, document, instrument, rule,
policy, observation, or institutional artifact, and it may preserve identifiers,
timestamps, provenance, and persistence relations among them.
Whether some event caused another, whether some decision was justified, or
whether some rule ought to apply are not commitments of the substrate merely
because the substrate identifies the relevant event, decision, rule, or
institutional artifact.

\begin{definition}[Referential Regime, from NS]
  \label{def:ns-referential-regime}
  A \emph{referential regime} is the triple of
  \begin{enumerate}
    \item individuation conditions,
    \item co-reference conditions, and
    \item persistence conditions
  \end{enumerate}
  by which a substrate fixes and tracks entities, occurrences, and institutional
  artifacts.
\end{definition}

Individuation conditions determine when something counts as one thing rather
than another.
Co-reference conditions determine when two references refer to the same thing.
Persistence conditions determine when something remains the same thing across
time, transformation, revision, or institutional change.
The referential regime is what allows a substrate to support shared use:
if parties cannot share enough reference to identify what is being interpreted,
they cannot reliably disagree about causes, norms, explanations, or obligations
concerning that thing.

\begin{definition}[Referential Commitments, from NS]
  \label{def:ns-referential-commitments}
  Let $\SubstrateRef$ denote the \emph{referential commitments} of $\Substrate$.
  These are the substrate-layer commitments fixed by its referential regime,
  including identifiers;
  the typing of entities, occurrences, and institutional artifacts;
  timestamps;
  provenance;
  and referential relations among them.
\end{definition}

\emph{Neutral Substrates} also defines permitted attribution propositions,
object-level causal or normative commitments, interpretive non-commitment,
extension stability, and neutrality by design.
Those definitions support its central \emph{constraint}:
under the stated contestability and common-ground assumptions,
a substrate's neutrality is guaranteed at design time when its
foundational layer is restricted
to referential commitments and permitted attribution propositions.
Specifically, the substrate must not make object-level causal or normative commitments.
Causal and normative content may appear through attributed assertions,
but the substrate commits to the attribution,
not to the asserted causal or normative proposition.

% -----------------------------------------------------------------------------
\subsection{Contribution}
\label{subsec:contribution}
% -----------------------------------------------------------------------------

\emph{Neutral Substrates} establishes the design constraint:
a neutral substrate requires stable reference while leaving causal and normative
interpretation to admissible extensions.
It does not analyze what makes stable reference achievable once the substrate's
records are transformed in ordinary use.
Records may be re-expressed, annotated, refined, decomposed, aggregated,
reassigned, substituted, forked, extended with provenance, or evolved through
state change.
Those operations place pressure on individuation, co-reference, and persistence.
This paper specializes the referential-regime component of that constraint
by deriving the lower-bound identity structure it imposes.

\begin{definition}[Transformation-Invariant Identity]
  \label{def:tinv}
  Within a neutral substrate, an \emph{identity regime} specifies how a
  referent's identity basis behaves under admissible transformations.
  For each transformation, the regime specifies whether the transformation
  preserves identity, breaks identity, or is identity-neutral.
  A referent is \emph{transformation-invariant} under transformations that do
  not break its declared identity basis.
\end{definition}

The dependency is:

\begin{center}
\begin{tabular}{@{}p{0.2\textwidth}>{\raggedright\arraybackslash}p{0.7\textwidth}@{}}
\textbf{Neutral Substrates:} &
a neutral substrate requires stable reference through individuation,
co-reference, and persistence.\\[2pt]
\textbf{This paper:} &
stable reference under ordinary transformations requires
transformation-invariant identity regimes, at least nine under the
assumptions fixed below.\\
\end{tabular}
\end{center}

Section~\ref{sec:algebra} makes Definition~\ref{def:tinv} operational as a
classification over a fixed transformation basis.
The rest of the paper computes the resulting lower bound.
