% !TeX root = se200_identity_regimes.tex
\section{Structural Requirements for Accountability Substrates}
\label{sec:requirements}

Given the assumptions and optimization criterion
fixed in Section~\ref{sec:assumptions},
the central question is not which ontology is most expressive or convenient,
but which structural distinctions are unavoidable.
We therefore state the representational requirements
that an ontological substrate must satisfy in order to
support accountability analysis under persistent disagreement.

These requirements are structural rather than interpretive.
They do not prescribe how accountability
should be understood, evaluated, or enforced; instead,
they specify what must be representable in a stable and neutral manner.

The requirements are framed as necessary capacities.
They are not derived from any
particular domain, policy area, or theory of governance,
but from the minimal conditions under which accountability claims
can be stated, inspected, and compared
without embedding causal or normative commitments in the substrate.

Before deriving the lower bound, we make explicit what is meant by an
\emph{identity-and-persistence regime}.

% Use asterisk to omit numbering from TOC this is just a definition
\definition*{Identity-and-Persistence Regime}
An identity-and-persistence regime specifies the conditions under which
entities remain the same entities over time and across incompatible extensions,
together with the relations they may admit without reclassification.

Two entities belong to the same regime if and only if
they share the same \stabilitycritical{} identity criteria,
persistence conditions, and admissible relations under
the neutrality and stability constraints adopted here.

Regimes are not introduced as ontological kinds or domain categories.
They are structural distinctions forced by the requirement that identity and
admissibility remain invariant under persistent disagreement.

The requirements enumerated below differ in scope.
R1 is a cross-cutting constraint:
it applies to every entity representable in the substrate,
regardless of which other requirements govern its individuation.
R2--R6 each specify representational capacities not entailed by R1 alone.

\subsection{R1: Stable Identity and Persistence (Global Constraint)}

\textbf{Requirement.}~The substrate must support stable reference to all entities
it represents over time.

Each entity must have identity and persistence conditions that are invariant
under incompatible extensions, independent of how the entity is later
classified, interpreted, or evaluated.
Changes in interpretation, role, status, or evaluation must not require
reclassification of entities or modification of their identity criteria.
Representations that determine identity contextually or retroactively
through interpretive layers fail to satisfy R1:
their identity conditions are not invariant under extension and
therefore do not constitute stable reference for accountability purposes.

In cross-institutional provenance, a research record transferred between
institutions must remain referable as the same record regardless of how
each institution classifies, annotates, or re-analyzes it.
If identity conditions shift when an institution introduces a new
classificatory scheme, the provenance chain breaks and accountability
for the record's content and custody cannot be maintained.
R1 is therefore non-optional wherever records must persist as stable
referents across frameworks that may individuate them differently.

\subsection{R2: Obligation-Bearing Capacity}

\textbf{Requirement.}~The substrate must support representation of entities that can bear
obligations or responsibilities independently of particular occurrences.

Such entities must be referable independently of the particular occurrences
in which obligations are fulfilled, violated, or discharged.
This requirement does not presuppose any substantive theory of normativity.
It requires only that obligation-bearing entities be representable as enduring
referents whose identity does not depend on specific events.

In multi-jurisdiction regulatory reporting, a corporate entity must be
referable as an obligation-bearer across legal regimes that may differ in
how they define corporate personhood, subsidiary responsibility, or
delegated liability.
If the substrate cannot represent obligation-bearing entities independently
of the specific events in which obligations are fulfilled or violated,
cross-jurisdiction audit comparisons require interpretive agreement on
the nature of the bearer itself, precisely the agreement that cannot
be assumed under persistent disagreement.
R2 is therefore non-optional wherever obligation-bearers must be
referable across legal or institutional frameworks that individuate
corporate personhood, subsidiary responsibility, or delegated
liability differently.

\subsection{R3: Normative Reference Without Execution}

\textbf{Requirement.}~The substrate must distinguish between the existence of authority or
obligation and the occurrences that execute, comply with, or violate it.

Normative or regulatory structures must therefore be representable
independently of the occurrences that enact, comply with, or violate them.
This distinction is necessary to support accountability over time, including
cases of delayed enforcement, partial compliance, amendment, or contested application.
Encoding normative reference solely through occurrences collapses authority
into execution and undermines stable reference to governing structures.

In legislative history tracking, a statutory authority must be referable
as a normative structure independently of the legislative acts that
enacted it, the administrative actions that applied it, and the judicial
decisions that interpreted it.
If normative reference is encoded solely through occurrences, a statute
that has been amended but not yet enforced, or contested but not yet
resolved, has no stable substrate-level referent.
R3 is therefore non-optional wherever accountability requires distinguishing
compliance with authority-as-written from
compliance with authority-as-executed, including cases of
amendment, delayed enforcement, or contested application.

\subsection{R4: Time-Indexed Occurrence}

\textbf{Requirement.}~The substrate must support representation of occurrences individuated
by temporal realization and provenance.

Occurrences must be individuated by temporal realization and provenance,
and must remain distinct from enduring entities.
This requirement applies to all discrete occurrences
relevant to accountability, including actions, filings, inspections, or transactions.
The list is illustrative; the requirement extends to any occurrence
for which temporal individuation and provenance are necessary for stable reference.
Occurrences are not required to carry interpretive or evaluative meaning; they
must assert only that something occurred at a particular time and place.

In multi-jurisdiction regulatory reporting, a filing, inspection, or
transaction must be individuated by when and where it occurred,
independently of how different jurisdictions classify or evaluate it.
Two jurisdictions may disagree on whether a given action constitutes
compliance, but both must be able to refer to the same occurrence as
their object of evaluation.
Without substrate-level individuation of time-indexed occurrences,
that shared reference is unavailable and cross-jurisdiction comparison
reduces to a dispute about what happened rather than
about how to evaluate it.
R4 is therefore non-optional wherever cross-jurisdiction accountability
requires shared reference to occurrences
independently of how those occurrences are classified or evaluated.

Requirements R2, R3, and R4 are mutually irreducible as representational capacities.
R2 requires stable reference to an enduring bearer;
R3 requires stable reference to a governing normative structure;
R4 requires stable reference to a discrete, time-indexed occurrence.
Satisfying any one does not entail satisfying the others.
None of the three capacities is definable in terms of the remaining two
without either circularity or the collapse of distinctions that
R1 requires to remain stable across incompatible extensions.

\subsection{R5: Applicability and Scope}

\textbf{Requirement.}~The substrate must support representation of applicability contexts
that scope where, to whom, or under what conditions a normative
structure applies.

It must be possible to state where, to whom, or under what conditions a
normative or regulatory structure applies, without collapsing such contexts
into either obligation-bearing entities or physical loci.
Applicability contexts may be nested, overlapping, or partially coincident.
They must therefore be representable as stable referents in their own right,
rather than as incidental attributes whose interpretation may vary across
extensions.

Applicability contexts cannot be reduced to properties of the R3 normative
structures they scope:
a single normative structure may admit different applicability contexts
under different jurisdictional frameworks, and
collapsing scope into the structure would require
its identity conditions to vary with those frameworks, violating R1.
Nor can applicability contexts be reduced to obligation-bearing entities (R2),
since scope specifies the domain over which authority applies rather than
identifying any particular bearer.

In multi-jurisdiction regulatory reporting, normative structures apply
to different sets of entities depending on jurisdiction, and those
jurisdictions may be nested, overlapping, or contested.
If applicability contexts cannot be represented as stable referents in
their own right, the scope of an obligation can only be recovered by
interpreting the obligation itself which reintroduces the shared
interpretive assumptions the substrate is designed to avoid.
R5 is non-optional wherever accountability claims must be scoped without
presupposing agreement on which normative framework or authority governs.

R2--R5 together establish the referential structure required for accountability analysis:
obligation-bearing entities (R2),
governing normative structures (R3),
time-indexed occurrences (R4),
and the applicability contexts that scope those structures (R5).
These capacities make it possible to represent
who is responsible,
which authority governs,
what occurred,
and where that authority applies.
They do not, however, establish stable reference to the descriptive records
or indicators associated with those entities and occurrences.

\subsection{R6: Descriptive Indicators Without Causal Commitment}

\textbf{Requirement.}~The substrate must support representation of descriptive indicators or
records that characterize states, performance, or outcomes associated
with entities or occurrences without asserting causal or normative
claims.

R6 is not entailed by R1--R5.
Stable reference to obligation-bearing entities (R2),
normative structures (R3),
time-indexed occurrences (R4),
and applicability contexts (R5)
does not establish stable reference to descriptive records associated with those referents.
Descriptive indicators have individuation conditions that are
independent of temporal realization (R4),
do not assert or constitute normative authority (R3),
do not identify obligation-bearers (R2),
and do not scope the application of normative structures (R5).
They must therefore be specified by a distinct requirement.

Such indicators must be representable without asserting causal relations,
evaluative judgments, or explanatory claims.
This requirement is necessary to support longitudinal comparison and outcome
inspection while preserving neutrality.
Collapsing indicators into occurrences or into interpretive annotations
eliminates stable reference to descriptive records
over time and across analytic frameworks.

In cross-institutional provenance, performance indicators and outcome
records must remain comparable across
institutions whose causal and evaluative frameworks differ,
for instance, where one institution models outcomes
in terms of intervention efficacy and another in terms of systemic risk factors.
If descriptive indicators are encoded as occurrences, they carry implicit
individuation conditions tied to causal assumptions; if encoded as
interpretive annotations, they are unstable under extension.
R6 is non-optional wherever longitudinal comparison of records must be
preserved without committing the substrate to any particular explanatory
or evaluative framework.

\subsection{Summary}

Together, R1--R6 specify the minimal representational capacities required for
an accountability substrate operating under persistent disagreement.
R1 is a cross-cutting stability constraint that applies to all
entities the substrate must represent.
R2--R5 specify four distinct representational capacities
that are mutually irreducible:
no one of them is entailed by or definable in terms of the others.
R6 specifies a further capacity not entailed by R1--R5.
These requirements do not impose a particular ontological partition;
they define the structural constraints under which any such partition must operate.
The following section uses R1--R6 to derive a lower bound
on the number of distinct identity-and-persistence regimes a substrate
must realize in order to satisfy them simultaneously.
