% !TeX root = se200_identity_regimes.tex
\section{Necessity: A Lower Bound on Ontological Structure}
\label{sec:necessity}

The previous section specified the structural requirements
for accountability substrates under neutrality and extension-stable identity.
We now use those requirements to derive a lower bound on the
individuation structure that any such substrate must realize.

The result is a necessity claim:
if a substrate is to support accountability under persistent disagreement
while maintaining neutrality and stable reference,
some minimum number of distinct identity-and-persistence regimes is unavoidable.
The lower bound is established by showing that certain representational
capacities cannot be realized within a single regime
without violating extension-invariant identity
or reintroducing hidden discriminators.

The regimes introduced below are persistence-and-admissibility profiles
induced when the requirements of Section~\ref{sec:requirements}
must be satisfied simultaneously under the neutrality and stability
constraints fixed in Section~\ref{sec:assumptions}.

% ============================================================
\subsection{Goal and Proof Strategy}
\label{sec:goal_proof_strategy}
% ============================================================

The goal of this section is to prove a numeric lower bound
on the number of identity-and-persistence regimes required
to satisfy R1--R6 simultaneously under the optimization criterion
fixed in Section~\ref{sec:assumptions}.

The argument has three steps.

First, we extract from R2--R6 (subject to R1) a set of
structurally distinct persistence-and-admissibility profiles
that must be realizable in any compliant substrate.
These profiles correspond to candidate identity-and-persistence regimes
forced by the requirements.

Second, we state a collapse admissibility criterion:
a proposed identification of two regimes is admissible only if
it preserves extension-invariant identity and admissible relations
without introducing interpretive flags, roles, or context-dependent
predicates at the extension layer.

Third, we show that any collapse among the forced profiles
violates the admissibility criterion through one of three failure modes:
identity instability, category error, or hidden regime reintroduction.
It follows that the substrate must realize at least as many distinct regimes
as there are forced profiles, establishing the lower bound.


% ============================================================
\subsection{Induced Identity-and-Persistence Regimes}
\label{sec:regimes}
% ============================================================

We now determine which identity-and-persistence regimes
are forced by R2--R6 when satisfied simultaneously
under R1's requirement of extension-invariant identity
and admissible relations.

An identity-and-persistence regime is characterized by
(i) persistence conditions and
(ii) the class of relations admissible without reclassification.

When the requirements are satisfied simultaneously under R1,
their persistence and admissibility constraints give rise to
distinct candidate identity-and-persistence profiles.
We extract these profiles in turn.

\subsubsection{Profile induced by R2: Obligation-bearing persistence}

R2 requires representation of entities that bear obligations independently of particular occurrences.
Such entities must persist across compliance, violation, delegation, and discharge.
Their admissible relations include normative attribution without altering identity.
This induces an initial distinct persistence-and-admissibility profile: obligation-bearing.

\subsubsection{Profile induced by R3: Normative-structure persistence}

R3 requires representation of normative or regulatory structures independently of their execution.
Such entities persist across enactment, amendment, interpretation, and repeal.
Their admissible relations include structural modification without collapsing into occurrences.
This induces a second distinct profile: normative-structure.

\subsubsection{Profile induced by R4: Time-indexed occurrence persistence}

R4 requires representation of occurrences individuated by temporal realization and provenance.
Such entities do not persist through modification but are fixed by their time-indexed realization.
This induces a third distinct persistence profile incompatible with enduring entities: time-indexed occurrence.

\subsubsection{Profile induced by R5: Applicability-context persistence}

R5 requires representation of scopes or applicability contexts as enduring entities.
Such entities admit nesting, overlap, and conditional scoping without collapsing into either authority or bearer entities.
This induces a fourth structurally distinct profile: applicability-context.

\subsubsection{Profile induced by R6: Descriptive-record persistence}

R6 requires representation of descriptive indicators without causal or normative commitment.
Such entities admit measurement and comparison relations without implying execution or authority.
This induces a fifth distinct profile: descriptive-record.

\subsubsection{Enduring Non-Normative Referents}

Satisfaction of R2--R5 presupposes stable reference to enduring
entities that are neither obligation-bearing nor authority-bearing,
yet may be regulated, inspected, or otherwise affected by occurrences.

Such entities persist independently of normative attribution
and independently of time-indexed realization.

Their persistence conditions are those of enduring referents
whose identity is not altered by normative attribution,
structural amendment, or temporal realization.

Their admissible relations include
participation in occurrences as patients or objects of action,
as well as measurement and description relations,
without implying normative authority or responsibility.

Profiles are individuated by the conjunction of
persistence conditions and admissible-relation classes.
Although ENR and REC both admit measurement and description relations,
their persistence conditions and participation roles differ,
yielding distinct identity-and-admissibility profiles.

This profile is not independently required by any single requirement.
However, stable reference to obligation-bearers,
normative structures, occurrences, and applicability contexts
requires referential anchoring to enduring entities
whose identity conditions and admissible relations
differ from each of those profiles.

This yields a sixth distinct persistence-and-admissibility profile:
enduring non-normative referents.

\subsubsection{Resulting Candidate Partition}

The induced profiles are pairwise non-equivalent under R1's stability constraint.
They differ either in persistence conditions or in admissible-relation classes.

Accordingly, any substrate satisfying R1--R6
must realize at least six distinct identity-and-persistence regimes.
For reference, we fix the following abbreviations
for these regimes:

\begin{enumerate}
\item \textbf{OBL: Obligation-bearing entity:} an enduring referent that may bear
      obligations or responsibilities and remain identifiable as a party across time.
\item \textbf{NOR: Normative structure:} an enduring referent that grounds
      obligations or permissions without being tied to any single occurrence.
\item \textbf{OCC: Time-indexed occurrence:} an entity individuated by its
      temporal realization and provenance.
\item \textbf{CTX: Applicability context:} a referent that scopes where, to whom,
      or under what conditions authority applies, and that may be nested or overlapping.
\item \textbf{REC: Descriptive record:} a referent that reports measured
      or derived properties without asserting causal or normative conclusions.
\item \textbf{ENR: Enduring non-normative referent:} an enduring entity
that may be regulated, inspected, or otherwise affected by occurrences,
but does not itself ground obligations or bear normative responsibility.
\end{enumerate}

% ============================================================
\subsection{Collapse Admissibility Criterion}
\label{sec:collapse_strategy}
% ============================================================

To establish a lower bound, we consider whether any two of the
induced identity-and-persistence regimes can be collapsed
into a single regime without violating the constraints of
Section~\ref{sec:assumptions}.

A collapse of two regimes is admissible only if the merged
representation realizes a single extension-stable identity
and persistence profile and supports a unified class of admissible relations
without introducing interpretive flags, role predicates,
or contextual discriminators at the extension layer,
which would reintroduce regime distinctions contrary to
the neutrality and extension-stability constraints of
Section~\ref{sec:assumptions}.

Under R1, identity and admissibility must be invariant across
incompatible extensions.
Accordingly, a collapse is inadmissible if it requires
any of the following:

\begin{itemize}
  \item \textbf{Identity instability:} the merged regime requires
        incompatible persistence criteria, so that identity judgments
        vary across extensions.
  \item \textbf{Category error:} the merged regime structurally admits
        relations inappropriate to some of its instances,
        requiring interpretive blocking rules.
  \item \textbf{Hidden regime reintroduction:} the original distinction
        is recovered through flags, roles, or contextual predicates,
        thereby reintroducing the regime at the extension layer.
\end{itemize}

Any collapse exhibiting one of these failure modes violates
either extension-invariant identity or the neutrality
criterion excluding hidden discriminators.

% ============================================================
\subsection{Pairwise Collapse Analysis}
\label{sec:collapse_examples}
% ============================================================

There are finitely many unordered pairs among the induced regimes.
We consider each class of collapse and show that
every such identification violates the admissibility criterion
by instantiating one of the three failure modes defined above.

\subsubsection{Category error: Obligation-bearing and Enduring non-normative referent}

Suppose an obligation-bearing entity and an enduring
non-normative referent are collapsed into a single regime.

The merged regime must admit a unified class of admissible relations.
Obligation-bearing entities admit normative attribution relations
such as \emph{hasObligation} or \emph{isResponsibleFor}
without reclassification.

If those relations are structurally admissible for all instances
of the merged regime, then every enduring referent becomes
eligible for normative attribution in principle,
embedding a normative commitment at the substrate level.

If, instead, normative attribution is restricted to only some instances,
implementations must introduce a discriminator
such as \emph{isObligationBearer} to block inappropriate attributions.
The original distinction is thereby reintroduced at the extension layer.

In either case, the collapse is inadmissible.

\subsubsection{Identity instability: Obligation-bearing and Normative structure}

Suppose an obligation-bearing entity and a normative structure
are collapsed into a single regime.

Obligation-bearing entities persist across compliance,
violation, and delegation.
Normative structures persist across enactment,
amendment, and repeal.

These persistence criteria are non-equivalent.
No single identity condition can preserve both
without making persistence judgments vary across extensions.

The collapse therefore violates extension-invariant identity under R1.

\subsubsection{Category error: Normative structure and Applicability context}

Suppose a normative structure and an applicability context
are collapsed into a single regime.

Normative structures admit structural modification relations
such as enactment, amendment, delegation, and repeal.
Applicability contexts instead admit scoping relations
such as nesting, overlap, and conditional refinement.

If the two are identified, the merged regime must admit
both modification-type and scoping-type relations
as structurally admissible.
However, these relation classes serve fundamentally different roles.
Amendment is a structural modification
that preserves the identity of a normative structure across change.
Whereas nesting and overlap are relations
among scopes that organize applicability conditions without modifying
normative content.

No single admissible-relation profile can unify
these relation classes.
Either the relations are admitted uniformly,
producing category error for some instances,
or their admissibility must be restricted contextually,
reintroducing the distinction through a hidden discriminator.

If all relations are admitted uniformly,
entities that are merely applicability scopes become amendable
normative structures, producing category error.

If relation admissibility is restricted by a predicate
such as \emph{isNormative} or \emph{isScope},
the original distinction is reintroduced
as a hidden regime at the extension layer.

The collapse is therefore inadmissible.

\subsubsection{Hidden regime reintroduction: Normative structure and Descriptive record}

Suppose a normative structure and a descriptive record
are collapsed into a single regime.

Normative structures ground obligations and admit
normative evaluation, enforcement, and modification.
Descriptive records report measured or derived properties
without asserting causal or normative commitments.

If both are identified, the merged regime must admit
a unified admissible-relation profile.
If normative evaluation relations are structurally admissible
for all instances of the merged regime,
then purely descriptive records become eligible
for normative grounding,
embedding evaluative commitments at the substrate level.

If, instead, normative evaluation is restricted to
a subset of instances,
implementations must introduce a discriminator
such as \emph{isNormative} or \emph{isDescriptive}
to distinguish grounding entities from descriptive ones.
Such a discriminator reintroduces the original distinction
at the extension layer,
violating the admissibility criterion and the neutrality constraint.

The collapse is therefore inadmissible.

\subsubsection{Summary of representative collapse failures.}

For completeness, we also note that collapsing time-indexed occurrences
with enduring non-normative referents produces identity instability,
since occurrence identity is fixed by temporal realization whereas
enduring referents persist independently of time-indexed realization.

Table~\ref{tab:collapse_summary} summarizes representative pairwise
collapses and the failure mode each induces under the admissibility
criterion defined in Section~\ref{sec:collapse_strategy}.

%\clearpage
\begin{table}[!htb]
\centering
\caption{Representative collapse failures under neutrality and stability constraints}
\label{tab:collapse_summary}
\vspace{1.0em}
\begin{tabular}{@{} p{3.0cm} p{3.5cm} p{5.2cm}@{} }
\hline
\textbf{Collapsed regimes} & \textbf{Failure mode} & \textbf{Reason for inadmissibility} \\
\hline
OBL / ENR
& Category error
& Normative attribution relations appropriate to obligation-bearing entities would become structurally admissible for all enduring referents, misclassifying non-normative entities as potential obligation-bearers. \\
\hline
OBL / NOR
& Identity instability
& Persistence as a responsible party and persistence as a normative structure require incompatible identity criteria across amendment and restructuring. \\
\hline
NOR / CTX
& Category error
& Modification relations (enactment, amendment, repeal) conflict structurally with scoping relations (nesting, overlap, conditional refinement); no unified admissible-relation profile can serve both
without misclassifying instances of one type as the other. \\
\hline
NOR / REC
& Hidden regime reintroduction
& Normative grounding and purely descriptive reporting admit incompatible admissibility profiles; implementations reintroduce \emph{isNormative} or \emph{isDescriptive} flags to recover the distinction. \\
\hline
OCC / ENR
& Identity instability
& Time-indexed individuation criteria conflict with extension-invariant persistence required for enduring entities. \\
\hline
\end{tabular}
\vspace{1.0em}
\end{table}

These three failure modes are exhaustive
relative to the admissibility criterion.
Since the admissibility criterion requires
exactly three conditions to hold jointly
(extension-invariant identity,
uniform admissibility of relations for the merged regime,
and neutrality with respect to hidden discriminators)
any collapse can violate admissibility only by
failing one of these conditions.
Identity instability corresponds to violation of the first,
category error to violation of the second,
and hidden regime reintroduction to violation of the third.

The remaining pairwise collapses instantiate the same three failure modes:
identity instability, category error, or hidden regime reintroduction.
Under the admissibility criterion of Section~\ref{sec:collapse_strategy},
no collapse eliminates a regime without violating neutrality or stable
reference (see Appendix~\ref{app:collapse} for the complete enumeration
of all fifteen pairs).

The following section establishes that this bound is tight:
a construction realizing exactly six regimes satisfies R1--R6
without hidden discriminators or additional identity categories,
confirming that six is the minimum achievable structure
under the stated constraints.

% ============================================================
\subsection{Lower-Bound Result}
% ============================================================

\begin{lemma}[Lower bound on identity-and-persistence regimes]
\label{lem:necessity}
Under the assumptions of Section~\ref{sec:assumptions},
any ontology satisfying requirements R1--R6
and the collapse admissibility criterion of
Section~\ref{sec:collapse_strategy}
must realize at least six distinct
identity-and-persistence regimes.
\end{lemma}

\begin{proof}[Proof sketch]
Section~\ref{sec:regimes} shows that satisfying R1--R6
under extension-invariant identity induces six
pairwise non-equivalent persistence-and-admissibility profiles.

Any ontology realizing fewer than six regimes must
identify  at least two of these profiles within a single regime.
By the collapse admissibility criterion
(Section~\ref{sec:collapse_strategy}),
every such identification produces either
identity instability, category error, or hidden regime reintroduction,
as demonstrated by exhaustive pairwise analysis (Section~\ref{sec:collapse_examples}
and Appendix~\ref{app:collapse}).

Each of these outcomes violates the neutrality
or stability constraints fixed in
Section~\ref{sec:assumptions}.

Therefore no ontology with fewer than six regimes
can satisfy R1--R6 simultaneously.
\end{proof}

% ============================================================
\subsection{Interpretation and Scope of the Necessity Result}
% ============================================================

The necessity result is conditional and structural.
It does not claim that fewer regimes are impossible in general, but that they
are incompatible with the optimization objective fixed in Section~\ref{sec:assumptions}.
Ontologies optimized for local agreement, negotiated semantics, or expressive
flexibility may legitimately employ fewer explicit distinctions by relying on
hidden or interpretive mechanisms.

Within the stated constraints, however,
the result establishes a minimal structural lower
bound: any neutral substrate capable of supporting accountability under
persistent disagreement must realize at least six distinct
identity-and-persistence regimes.
The following section shows that this bound is tight by exhibiting a
construction that satisfies all requirements with a minimal set of six regimes.
