% !TeX root = ../se200_identity_regimes.tex

% =============================================================================
\section{Reference Inventory}
\label{sec:inventory}
% =============================================================================

Before asking how identity behaves, we fix what there is to individuate.
An accountability substrate must let one state, inspect, and
compare claims about who is responsible, which authority governs, what happened, where it applies, what was recorded, and what was acted upon.
It must do this without embedding causal or normative content at the substrate level.
That imposes a short inventory of reference kinds.
The inventory is modest:
not a theorem, not an upper ontology,
and not a claim that these are the only things a domain may represent.
It states that an accountability substrate under the assumptions of
Section~\ref{sec:formal_setting}
must be able to refer to these kinds of things.

The inventory supplies identity carriers.
The transformation algebra of Section~\ref{sec:algebra} determines how those carriers are individuated.
This distinction matters.
A reference kind says what must remain addressable.
An identity regime says what counts as the same referent under transformation.
The central result of the paper is that the number of identity regimes
required by stable reference exceeds the number of reference kinds in the inventory.

% -----------------------------------------------------------------------------
\subsection{Reference Kinds and Identity Carriers}
\label{subsec:reference_kinds_identity_carriers}
% -----------------------------------------------------------------------------

\begin{definition}[Reference kind]
\label{def:reference_kind}
A \emph{reference kind} is a class of referents that a
neutral accountability substrate must be able to identify, track, and
distinguish from other referents without making object-level causal or normative commitments.
\end{definition}

\begin{definition}[Identity carrier]
\label{def:identity_carrier}
An \emph{identity carrier} is a referential position in the substrate to
which an identity regime may be assigned.
It is the object whose sameness, difference, and persistence are evaluated under transformation.
\end{definition}

Reference kinds supply identity carriers.
They do not themselves determine identity regimes.
A single reference kind may support more than one identity regime when
different identity bases respond differently to the same transformation.
Conversely, the fact that two carriers are both represented in the
substrate does not imply that they share an identity regime.
Regime assignment depends on transformation behavior, not merely on the name of the kind.

The inventory used in this paper is:

\begin{center}
  \begin{tabular}{@{}l>{\raggedright\arraybackslash}p{0.68\textwidth}@{}}
    \toprule
    Reference kind & What must remain referable \\
    \midrule
    Obligation-bearer &
    a party that can bear responsibility across the occurrences
    discharging, violating, or contesting that responsibility \\
    Rule &
    a governing normative structure, distinct from the acts that enact, amend, apply, or challenge it \\
    Occurrence &
    a happening individuated by temporal realization and provenance \\
    Scope &
    the domain where, to whom, or under what conditions a rule applies \\
    Record &
    a descriptive indicator or record that carries reference without
    asserting causal or normative truth \\
    Plain referent &
    a thing acted upon, regulated, located, transferred, or observed
    that neither bears obligations nor grounds authority merely as such \\
    \bottomrule
  \end{tabular}
\end{center}

These six reference kinds are familiar from foundational
ontology, legal ontology, provenance, and institutional recordkeeping.
The paper does not claim novelty for the inventory itself.
The novelty lies in showing that this inventory does not settle the
identity structure required by a neutral substrate.
Three of the six kinds split when ordinary transformations are considered.

% -----------------------------------------------------------------------------
\subsection{Obligation-Bearers}
\label{subsec:inventory_obligation_bearers}
% -----------------------------------------------------------------------------

An accountability substrate must support reference to obligation-bearing entities.
An obligation-bearer is a party that can bear
responsibility, duty, liability, authority, or accountability
across occurrences in which those statuses are fulfilled, violated, transferred, contested, or discharged.
The substrate need not decide whether a party is actually responsible in a particular case.
It must only preserve reference to the party that a framework may treat as responsible.

In multi-jurisdiction regulatory reporting, a corporate entity may be treated differently across jurisdictions.
One jurisdiction may individuate responsibility at the parent-entity level.
Another may individuate it at the subsidiary, facility, role, or delegated-agent level.
A neutral substrate need not settle which legal theory governs.
It must still allow the relevant bearer to remain referable across those competing frameworks.

Obligation-bearers cannot be reduced to the occurrences in which obligations are discharged.
The same bearer may participate in many occurrences.
The same occurrence may involve several bearers.
Collapsing bearer into occurrence destroys the ability to compare responsibility across time and across frameworks.

% -----------------------------------------------------------------------------
\subsection{Rules}
\label{subsec:inventory_rules}
% -----------------------------------------------------------------------------

An accountability substrate must support reference to governing rules or normative structures.
A rule is a structure that may be enacted, amended, delegated, interpreted, applied, violated, or repealed.
The substrate need not assert that the rule is valid, justified, binding, or correctly interpreted.
It must preserve reference to the rule as the object of such attributed claims.

In legislative history tracking, a statute must remain referable
independently of the legislative act that enacted it, the amendment
that altered it, the agency action that applied it, and the judicial decision that interpreted it.
A statute can be contested while still remaining the same referent
for purposes of citation, comparison, and provenance.
If the rule is collapsed into its enactment or application occurrence,
then later disagreement about the occurrence changes reference to the rule itself.

Rules therefore supply identity carriers distinct from occurrences.
They also supply one of the important split cases below.
A rule may be individuated by what it requires or by how its text or internal structure is organized.
That difference is not settled by naming the kind ``rule.''
It is settled by transformation behavior.

% -----------------------------------------------------------------------------
\subsection{Occurrences}
\label{subsec:inventory_occurrences}
% -----------------------------------------------------------------------------

An accountability substrate must support reference to occurrences.
An occurrence is a happening individuated by temporal realization and provenance.
Examples include filings, inspections, votes, transfers, transactions,
observations, decisions, enactments, amendments, and enforcement actions.
The list is illustrative.
What matters is that the occurrence can be referred to as something that happened,
independently of how competing frameworks evaluate it.

In multi-jurisdiction regulatory reporting, two jurisdictions may disagree about
whether a filing constitutes compliance.
They must nevertheless be able to refer to the same filing.
In legislative history tracking, parties may disagree about the legal significance of an enactment.
They must nevertheless be able to refer to the same enactment occurrence.
In provenance analysis, institutions may disagree about the meaning or reliability of a measurement.
They must nevertheless be able to refer to the same measurement occurrence.

Occurrences cannot be reduced to obligation-bearers, rules, records, or scopes.
They are time-indexed referents.
They may instantiate, affect, or be described by other carriers, but they are not identical with them.

% -----------------------------------------------------------------------------
\subsection{Scopes}
\label{subsec:inventory_scopes}
% -----------------------------------------------------------------------------

An accountability substrate must support reference to scopes.
A scope is the domain where, to whom, or under what conditions a rule applies.
Scopes may be territorial, institutional, temporal, procedural, classificatory, jurisdictional, or conditional.
They may be nested, overlapping, partially coincident, or contested.

In multi-jurisdiction regulatory reporting, the same rule may apply to
different entities, places, facilities, transactions, or reporting periods under different frameworks.
If scope is represented only as an attribute of the rule, then a change in
applicability can appear to change the identity of the rule.
If scope is represented only as a property of obligation-bearers, then the
domain of applicability cannot be compared independently of the parties within it.
A neutral substrate needs scope as a referent in its own right.

Scopes supply another split case.
A scope may be individuated extensionally, by what it covers.
It may also be individuated structurally, by how its internal jurisdictional or applicability structure is organized.
Decomposing a scope can preserve its extension while changing its structure.
The reference kind ``scope'' therefore does not determine a single identity regime.

% -----------------------------------------------------------------------------
\subsection{Records}
\label{subsec:inventory_records}
% -----------------------------------------------------------------------------

An accountability substrate must support reference to descriptive records.
A record is a descriptive indicator, document, dataset entry, observation record,
provenance item, or other referent that can be cited, transferred, compared, annotated, or analyzed without the substrate asserting a causal or normative claim.
The substrate may record that a source asserts a claim.
It does not thereby assert the claim.

In cross-institutional provenance, a record may be produced by
one institution, transferred to another, annotated by a third, and re-analyzed by a fourth.
The institutions may disagree about what the record shows or whether it supports a conclusion.
They must nevertheless be able to refer to the same record through the chain of custody, annotation, and reuse.

Records cannot be reduced to the entities or occurrences they describe.
A record may describe an occurrence without being that occurrence.
A record may concern a rule without being that rule.
A record may support an attributed causal or normative claim without
making that claim at the substrate level.
Stable reference to records is therefore required independently of the referents described by records.

% -----------------------------------------------------------------------------
\subsection{Plain Referents}
\label{subsec:inventory_plain_referents}
% -----------------------------------------------------------------------------

An accountability substrate must also support reference to plain referents.
A plain referent is a thing acted upon, regulated, located, transferred, observed,
measured, or described that neither bears obligations nor grounds authority merely as such.
Examples include facilities, parcels, instruments, devices, environmental sites,
physical objects, infrastructure components, samples, and assets.
The category is intentionally modest.
It captures referents that are important to accountability without already being
obligation-bearers, rules, occurrences, scopes, or records.

Plain referents are needed because many accountability questions concern things
that are acted upon or regulated.
A regulated facility may be inspected.
A parcel may fall within a jurisdictional scope.
An instrument may produce a measurement record.
A sample may be transferred across institutions.
The substrate must preserve reference to the thing without deciding which
causal or normative interpretation applies to it.

Plain referents supply the third split case.
A referent may be individuated by locus, such as the site or place that remains fixed.
It may also be individuated by the object occupying that locus.
Forking or replacement can preserve the locus while changing the object, or preserve
the object while changing the locus.
The reference kind ``plain referent'' therefore does not determine a single identity regime.

% -----------------------------------------------------------------------------
\subsection{Summary}
\label{subsec:inventory_summary}
% -----------------------------------------------------------------------------

Counting reference kinds gives six:
obligation-bearer, rule, occurrence, scope, record, and plain referent.
Counting identity regimes will yield nine (Section~\ref{sec:nine}).
The inventory does not derive that lower bound.
It supplies the carriers on which the bound operates once the transformation
basis is applied, which is why the paper distinguishes reference kinds from
identity regimes.
The surplus regimes appear because three reference kinds underdetermine their identity basis.
Rules split into content-fixed and structure-fixed regimes.
Scopes split into extension-fixed and structure-fixed regimes.
Plain referents split into locus-fixed and object-fixed regimes.

This distinction addresses the main methodological risk.
An argument that listed six requirements and then named six regimes would be circular.
Here the inventory names only what must remain referable, the transformation
algebra determines how identity behaves, and the nine-regime lower-bound core follows
from the mismatch between those two levels.

The next section applies the transformation basis of
Section~\ref{sec:formal_setting} to these six carriers, determining which can
share identity behavior and which must split.
