% !TeX root = se200_identity_regimes.tex
\section{Sufficiency: A Six-Regime Construction}
\label{sec:sufficiency}

\subsection{The Six Identity-and-Persistence Regimes}

This section exhibits a construction realizing
the six identity-and-persistence regimes
derived in Section~\ref{sec:necessity}.
No additional regimes are introduced,
and none of the induced regimes are merged or weakened.

Each regime is implemented with the identity conditions
and admissible-relation profile specified in the derivation.
The construction fixes these regimes as pairwise disjoint,
with extension-invariant persistence criteria,
and without interpretive flags or role-based discriminators.

We now specify a minimal set of typed relations
that connects these regimes while preserving
neutrality and stable reference.

\subsection{Admissible Relations}

The construction introduces the following typed relations.
Together these relations realize the admissibility profiles
required to satisfy R2--R6.
Each requirement is realized by at least one listed relation.
The coverage mapping is made explicit in
Section~\ref{sec:requirements_verification}.

\begin{itemize}

\item authority grounding and delegation:
{\small
\[
\begin{array}{lcl}
\textit{enacts},\ \textit{issues} &:& \mathrm{OBL} \to \mathrm{NOR}
\end{array}
\]
}

\item participation and party status:
{\small
\[
\begin{array}{lcl}
\textit{party\text{-}to}    &:& \mathrm{OBL} \to \mathrm{NOR} \\
\textit{involves}           &:& \mathrm{OCC} \to \mathrm{OBL} \\
\textit{has\text{-}patient} &:& \mathrm{OCC} \to \mathrm{ENR}
\end{array}
\]
}

\item execution under authority:
{\small
\[
\begin{array}{lcl}
\textit{occurs\text{-}under} &:& \mathrm{OCC} \to \mathrm{NOR}
\end{array}
\]
}

\item applicability:
{\small
\[
\begin{array}{lcl}
\textit{applies\text{-}in} &:& \mathrm{NOR} \to \mathrm{CTX} \\
\textit{refines}           &:& \mathrm{CTX} \to \mathrm{CTX} \\
\textit{overlaps}          &:& \mathrm{CTX} \leftrightarrow \mathrm{CTX}
\end{array}
\]
}

\item description:
{\small
\[
\begin{array}{lcl}
\textit{measures} &:& \mathrm{REC} \to (\mathrm{OBL} \mid \mathrm{ENR} \mid \mathrm{CTX})
\end{array}
\]
}

\end{itemize}

Relations in this construction are structural and descriptive.
They assert reference, participation, scoping, or provenance,
but do not encode causal mechanism or normative validity.

Authority grounding and delegation records that an
obligation-bearing entity enacts or issues a normative structure,
without asserting validity or enforceability.

Participation relations record that an occurrence
is associated with obligation-bearing entities via \emph{involves}
and with enduring non-normative referents via \emph{has-patient}.

The \emph{measures} relation associates a descriptive record
with an enduring referent,
without asserting accuracy, causal relevance,
or evaluative significance.

Occurrences serve as temporal and provenance anchors
for participation and measurement relations.
Descriptive records target enduring referents whose identity
persists across incompatible extensions.

This relation set suffices to satisfy R1--R6
without introducing additional identity regimes,
role-based typing, or extension-layer discriminators.

Temporal and provenance information attach only to OCC and REC
via typed attributes or admissible relations.
These annotations do not introduce new identity criteria
or additional identity-bearing categories.

\subsection{Satisfaction of Structural Requirements}
\label{sec:requirements_verification}

The relation set defined above satisfies R1--R6 by the
following coverage mapping:

\begin{itemize}
  \item \textbf{R1 (Stable identity):}
        Each regime fixes an extension-invariant
        identity-and-persistence criterion.
        No relation in the construction propagates
        persistence criteria across regime boundaries.
  \item \textbf{R2 (Obligation-bearing capacity):}
        \textit{enacts}, \textit{issues}, \textit{party-to},
        and \textit{involves} realize normative attribution
        for OBL without reclassification.
  \item \textbf{R3 (Normative reference without execution):}
        \textit{enacts}, \textit{issues}, and \textit{applies-in}
        ground NOR independently of any OCC instance.
  \item \textbf{R4 (Time-indexed occurrence):}
        \textit{involves}, \textit{has-patient}, and
        \textit{occurs-under} anchor OCC to its participants
        and governing normative structure
        without altering their identity.
  \item \textbf{R5 (Applicability and scope):}
        \textit{applies-in} scopes NOR within CTX;
        \textit{refines} realizes conditional nesting;
        \textit{overlaps} realizes non-hierarchical
        partial scope without requiring containment.
  \item \textbf{R6 (Descriptive indicators):}
        \textit{measures} associates REC with enduring
        referents (OBL, ENR, CTX) without asserting
        causal or normative conclusions.
\end{itemize}

No requirement introduces a relation that propagates
persistence criteria or normative commitments
outside the regime for which it is specified.
No additional identity regimes or extension-layer
discriminators are required.

\subsection{Sufficiency Proposition}

\begin{proposition}[Sufficiency of six regimes]
\label{prop:sufficiency-six}
The six-regime construction specified in this section
satisfies R1--R6 under the assumptions of
Section~\ref{sec:assumptions}.
\end{proposition}

\begin{proof}
By construction, each of R2--R6 is realized
by at least one typed relation
in the construction,
and all regimes satisfy R1.
The regimes are pairwise disjoint,
and the typed relation set connects them
only through descriptive, participation, and scoping relations
whose admissibility conditions
do not propagate persistence criteria across regime boundaries.
and the construction introduces no interpretive flags,
role-based discriminators, or additional identity regimes.
Hence the construction satisfies R1--R6.
\end{proof}

\subsection{Tightness of the Bound}

Lemma~\ref{lem:necessity} establishes that any ontology satisfying R1--R6
must realize at least six distinct identity-and-persistence regimes.
Proposition~\ref{prop:sufficiency-six} shows that the construction
presented in this section satisfies those requirements with six regimes.

The six regimes of the construction are pairwise non-equivalent
under R1's stability constraint:
each pair differs in persistence conditions,
admissible-relation classes,
or both,
as established in Section~\ref{sec:regimes}.
Any identification of two regimes would therefore instantiate one of
the failure modes defined in Section~\ref{sec:collapse_strategy},
violating either extension-invariant identity or the neutrality constraint.

Thus the lower bound established in Lemma~\ref{lem:necessity}
is achieved exactly by the present construction.
Six regimes are both necessary and sufficient,
and the bound is therefore tight.

No compliant construction can realize fewer regimes
without violating either extension-invariant identity
or the neutrality constraint.
