% !TeX root = se200_identity_regimes.tex
\section{Interpretation of the Result}
\label{sec:interpretation}

This section clarifies how the lower- and upper-bound results should be
understood, and how they should not be understood.
The aim is not to defend the result against alternative modeling goals,
but to situate it relative to the explicit optimization criterion fixed
earlier in the paper.

\subsection{What the Result Establishes}

Taken together, the necessity and sufficiency results establish a
conditional optimality claim.
Under the assumptions of Section~\ref{sec:assumptions},
any ontology intended to function as a neutral substrate
for accountability under persistent disagreement
must realize at least six distinct identity-and-persistence regimes.
Conversely, a construction with a minimal set of six such regimes suffices to meet
all stated representational requirements without introducing hidden
structure or interpretive commitments.

The result is therefore tight.
It does not merely show adequacy or inadequacy
under particular conditions;
it establishes that six regimes are both necessary and sufficient
under the stated optimization objective.

\begin{theorem}[Tightness of the six-regime bound]
\label{thm:tightness}
Under the assumptions of Sections~\ref{sec:assumptions}--\ref{sec:requirements},
exactly six identity-and-persistence regimes are necessary and sufficient
for a neutral accountability substrate under persistent disagreement.
\end{theorem}
\begin{proof}
The theorem follows directly from Lemma~\ref{lem:necessity}
and Proposition~\ref{prop:sufficiency-six}.
\end{proof}

\subsection{Minimality Relative to Stability}

The result should not be read as a claim about simplicity in an intuitive
or engineering sense.
Ontologies with fewer explicit entity kinds may be easier to describe or extend
within a coordinated interpretive community.
However, when such reductions are achieved by encoding \stabilitycritical{}
distinctions as roles, flags, or contextual predicates, total structural
complexity is not reduced.
Instead, it is displaced into extension layers whose interpretation cannot
be held fixed under disagreement.

In this paper, minimality is defined relative to stability under persistent disagreement.
\emph{A representation is minimal only if removing explicit structure does not
require reintroducing the same distinctions implicitly.}
From this perspective, an ontology with fewer explicit kinds but greater
interpretive burden is not simpler, but less explicit.

\subsection{Why Hidden Regimes Are Excluded}

Hidden regimes are often effective where interpretation is shared,
enforced, or low-cost to renegotiate.
They support flexibility and expressive economy within a single
governance framework.

The present analysis excludes such regimes not because they are incorrect,
but because they are incompatible with the stated objective.
When disagreement is persistent and uncoordinated, distinctions that are
not fixed at the substrate level cannot be relied upon to preserve identity
or admissible relations.
Stability then depends on interpretive convergence, which the substrate is
explicitly designed not to assume.

\subsection{Relation to Other Ontological Strategies}

Many upper ontologies and modeling frameworks adopt different optimization
objectives, including metaphysical coverage, expressive richness, or
alignment with natural language or scientific theory.
Such frameworks may introduce additional kinds or tolerate contextual
typing precisely in order to support richer modeling.

The result presented here neither subsumes nor contradicts those frameworks.
It identifies a specific point in the design space:
the minimal structure required for a neutral, extension-stable substrate under persistent disagreement.
Frameworks optimized for other goals may legitimately occupy different
points in that space.

The following comparison is illustrative rather than evaluative.
Table~\ref{tab:regime-mapping} shows common representation
strategies for these regimes in BFO~2.0 and DOLCE and highlights
where regime-level distinctions required for neutrality are not
structurally enforced.

The comparison is structural rather than taxonomic;
it identifies where regime-level distinctions
required for neutrality are left to implementation
rather than enforced at the type level.

The table identifies the structural under-constraint that would arise in each case.
These under-constraints are not problems in their intended use.
They are points where neutral substrate implementations built on those frameworks
might need to introduce additional modeling commitments
that the present substrate treats as regime-level structure.

The six regimes are not derived from gaps in existing frameworks.
Rather, independently motivated requirements reveal where those frameworks
leave regime-level distinctions to the implementer.

%\clearpage
\begin{table}[!htbp]
\centering
\caption{Illustrative representation strategies for the six regimes
in BFO~2.0 and DOLCE, with structural under-constraints
under neutrality and stability requirements.}
\label{tab:regime-mapping}
\small
\renewcommand{\arraystretch}{1.15}
\begin{tabular}{@{} p{1.5cm} p{5.2cm} p{4.7cm}@{}}
\toprule
\textbf{Regime} & \textbf{BFO 2.0} & \textbf{DOLCE} \\
\midrule

ENR (enduring non-normative referent) & Independent Continuant (material object).
Clean for material entities; institutional and fiat entities
require additional modeling choices not enforced by the framework. &
Physical or Social Object (endurant).
Social object is first-class, providing a cleaner fit for
institutional referents than BFO. \\

OBL (obligation-bearing entity) &
Specifically Dependent Continuant (Role or Disposition)
inhering in an Independent Continuant.
Obligation-bearing is not a first-class category;
it is realized via a role--bearer pair,
splitting what the present substrate treats as a single regime. &
Quality or property of a Social Object.
Obligation-bearing capacity is not a structural first-class
distinction; the OBL/NOR boundary remains implementation-dependent. \\

NOR (normative structure) &
Generically Dependent Continuant (information artifact,
normative document).
The substrate-level separation between authority-as-referent (NOR)
and authority-as-executed (OCC) is a modeling choice,
not structurally enforced. &
Social Object (normative structure).
Better motivated than BFO, but authority and execution
can both be represented as social objects or perdurants
without a regime-level firewall between NOR and OCC. \\

OCC (time-indexed occurrence) &
Process or Process Boundary (occurrent).
Closest mapping in both frameworks;
BFO occurrents are well-individuated temporally. &
Perdurant (Event or Process).
DOLCE distinguishes achievements from activities,
which is richer than required but not incompatible. \\

CTX (applicability context) &
Site (immaterial Independent Continuant) for spatial scope only.
Non-spatial applicability contexts (normative jurisdiction,
conditional scope) have no stable BFO home and collapse
into GDC, conflating CTX with NOR. &
Abstract Region or Feature.
DOLCE regions are spatiotemporal;
normative applicability collapses into Social Object
(conflating CTX with NOR) or Quality (conflating CTX with REC). \\

REC (descriptive indicators without causal commitment) &
Generically Dependent Continuant (data item,
measurement information entity).
Both normative structures (NOR) and descriptive indicators (REC)
map to GDC; the stability-critical distinction between them
is not explicitly represented at the BFO type level. &
Quality (with associated quale or region).
DOLCE's quality theory is the strongest match;
the without-causal-commitment constraint is a modeling
discipline, however, not structurally enforced. \\

\bottomrule
\end{tabular}
\end{table}

Table~\ref{tab:underconstraints} summarizes the systematic under-constraints
that arise in both frameworks relative to the neutrality and stability requirements.

%\clearpage
\begin{table}[!hb]
\centering
\caption{Systematic under-constraints in BFO~2.0 and DOLCE relative to
neutrality and stability requirements.}
\label{tab:underconstraints}
\small
\renewcommand{\arraystretch}{1.35}
\begin{tabular}{@{} p{2cm} p{4.7cm} p{4.7cm} @{}}
\toprule
\textbf{Under-constraint} & \textbf{BFO 2.0} & \textbf{DOLCE} \\
\midrule


NOR/REC conflated &
Both normative structures and descriptive indicators map to GDC;
the stability-critical distinction between them is invisible
at the type level and left to the implementer. &
Both map to Social Object or Quality depending on implementation;
the distinction is a modeling discipline, not structurally enforced. \\

CTX has no structural home &
Non-spatial applicability contexts are represented as GDC,
conflating CTX with NOR. &
Normative applicability collapses into Social Object (conflating CTX
with NOR) or Quality (conflating CTX with REC). \\

OBL/NOR boundary implementation-dependent &
The separation between obligation-bearer and normative structure
is mediated by the inheres-in relation, not enforced by
regime membership; unstable under delegated authority. &
The same boundary is mediated by participation relations;
regime-level enforcement is absent. \\

\bottomrule
\end{tabular}
\smallskip
\begin{minipage}{\textwidth}
\footnotesize
\textit{Implementations that resolve these under-constraints silently
introduce hidden regimes in exactly the sense of
Section~\ref{sec:necessity}.}
\end{minipage}
\end{table}

\subsection{Conditional Scope of the Claim}

The lower-bound result is explicitly conditional.
If any assumption of Section~\ref{sec:assumptions} is rejected,
such as neutrality, stability under incompatible extension,
or the exclusion of hidden regimes,
the optimization criterion is invalidated
and the bound no longer applies.
We do not claim that these assumptions are universally correct;
we claim only that the result follows when they are adopted.

If stability is weakened to allow context-dependent
regime membership (e.g., role- or viewpoint-relative identity/admissibility),
some regimes may be representable as
role- or parameter-dependent refinements of fewer base regimes,
but doing so shifts structural burden into
coordination, versioning, or exchange mechanisms
whose admissibility conditions are not fixed
at the substrate level and are outside the scope of this analysis;
the lower bound therefore does not apply under such weakenings.

Within the stated scope, however, the result is unavoidable.
Under the stated assumptions, any attempt to reduce the number of regimes
would require sacrificing stability,
reintroducing hidden distinctions, or weakening the neutrality constraint.
The six-regime structure thus emerges not as a design preference, but as a
structural consequence of the problem as posed.

The present analysis does not model state transitions,
versioning mechanisms, or exchange protocols.
Such structures may be required for practical interoperability,
but they operate downstream of the regime-level identity constraints
established here.

\subsection{Implications}

The primary implication of this result is diagnostic rather than
prescriptive.
It provides a principled way to evaluate claims that an ontology supports
neutral interoperability under disagreement.
If such a framework relies on implicit roles or context-dependent typing
for \stabilitycritical{} distinctions, the present result explains why those
claims cannot be sustained simultaneously under the stated constraints.

Conversely, for designers who share the stated optimization objective, the
result identifies which distinctions must be made explicit,
and which may safely be deferred to extension layers,
in order to preserve neutrality and stable reference
across incompatible interpretations.
The six-regime structure thus serves both as a lower bound and
as a design template for accountability substrates in contested domains.

% Address: how to axiomatize certain relationships and entities
% so as to leave room for subsequent, potentially divergent, specializations.
The resulting regime structure can be understood
as an axiomatization pattern for substrates
intended to support divergent specialization:
identity and admissibility constraints
are fixed at the substrate level so that
subsequent causal or normative extensions
may diverge without retroactive reinterpretation.

% REVISE TO FINISH THREADING THE EXAMPLE CASES

\subsection{Grounding the Construction in the Running Cases}

The six-regime construction is not an abstract partition.
Each case introduced in Section~\ref{sec:intro} exercises the full structure and
illustrates what stable reference under persistent disagreement concretely requires.

\subsubsection{Multi-jurisdiction regulatory reporting.}
A corporate entity is represented as an obligation-bearing referent (OBL)
that persists across jurisdictional differences in how corporate
personhood or subsidiary liability is defined.
The reporting statute under which it operates is represented as an
authority-bearing structure (NOR) that persists across amendments,
contested interpretations, and delayed enforcement.
Each filing, inspection, or transaction is represented as a
time-indexed occurrence (OCC) individuated by when and where it happened,
independently of how any jurisdiction evaluates it.
The jurisdiction itself is represented as an applicability context (CTX)
that scopes which obligations apply to which entities without collapsing
into either the statute or the filer.
Performance and compliance indicators are represented as descriptive
records (REC) that remain comparable across jurisdictions without
carrying evaluative or causal commitments.
Any entity that is regulated but does not itself bear obligations,
such as a regulated substance or a monitored facility, is
represented as an acted-upon referent (ENR).
The six-regime construction allows two jurisdictions to refer to the
same corporation, the same statute, and the same filing while
maintaining incompatible evaluations of compliance,
without either forcing interpretive convergence or losing
the shared referential ground required for audit comparison.

\subsubsection{Legislative history tracking.}
A statute is represented as an authority-bearing structure (NOR) that
remains the same normative referent across the legislative acts that
enacted it (OCC), the amendments that modified it (OCC), and the
administrative decisions that applied it (OCC).
The legislature and the agencies that implement the statute are
represented as obligation-bearing referents (OBL) whose identity
does not depend on any particular enactment event.
Jurisdictions and conditional scopes such as the geographic or
subject-matter boundaries of the statute's application, are
represented as applicability contexts (CTX) that may be nested,
overlapping, or contested without destabilizing reference to the
statute itself.
Records of legislative proceedings, judicial interpretations, and
compliance histories are represented as descriptive indicators (REC)
that characterize outcomes without asserting which interpretation
of the statute is correct.
The construction allows legal historians, policy analysts, and
regulators working under incompatible interpretive frameworks to
refer to the same statutory authority and the same sequence of
enactment events while disagreeing entirely on what the statute
means or which application was legitimate.

\subsubsection{Cross-institutional provenance.}
A research or administrative record is represented as an
acted-upon referent (ENR) whose identity and persistence conditions
are invariant under the classificatory schemes of any receiving
institution.
The institution that produced the record and the institutions that
subsequently hold or analyze it are represented as obligation-bearing
referents (OBL) that remain identifiable as custodians across
transfers, mergers, and re-organizations.
Policies governing data use, retention, or sharing are represented
as authority-bearing structures (NOR) that persist independently of
whether they have been enforced or contested in any particular transfer.
Each transfer, annotation event, or re-analysis is represented as a
time-indexed occurrence (OCC) individuated by its provenance,
independently of how institutions classify the action.
Institutional scope conditions, such as which data a policy governs
or which entities fall under a data-sharing agreement, are
represented as applicability contexts (CTX).
Derived indicators, quality metrics, and outcome summaries attached
to the record are represented as descriptive records (REC) that remain
comparable across institutions whose causal and evaluative frameworks
differ.
The construction preserves the provenance chain as a stable referential
structure even when no two institutions agree on how the record should
be classified, evaluated, or interpreted.
