% !TeX root = ../se200_referential_regimes.tex

% =============================================================================
\section{The Lower Bound}
\label{sec:bound}
% =============================================================================

The previous section collected the nine core identity regimes used in this paper.
This section proves the lower-bound claim.
The proof does not require showing that these nine regimes exhaust all admissible identity bases.
It requires only two facts:
the inventory requires stable reference to six carrier kinds,
and three of those carrier kinds exhibit witnessed split pressure under the transformation basis.

% -----------------------------------------------------------------------------
\subsection{The Inventory Floor}
\label{subsec:inventory_floor}
% -----------------------------------------------------------------------------

The reference inventory of Section~\ref{sec:inventory} requires stable reference to six kinds of carrier:
obligation-bearers, rules, occurrences, scopes, records, and plain referents.
Stable reference to a carrier kind requires an identity regime for that kind, because the substrate must determine when representations of that carrier count as the same referent.

\begin{lemma}[Inventory floor]
\label{se200.lem.InventoryFloor}
Under the assumptions of Section~\ref{sec:formal_setting},
a neutral accountability substrate realizing the inventory of Section~\ref{sec:inventory}
must realize at least one identity regime for each required carrier kind.
\end{lemma}

\begin{proof}
By Definition~\ref{def:reference_kind}, each reference kind in the inventory
must be identifiable, trackable, and distinguishable from other referents
without making object-level causal or normative commitments.
By Definition~\ref{def:identity_regime}, an identity regime states what counts as
sameness for a referent under transformation.
Thus each required carrier kind needs identity conditions sufficient to
preserve stable reference to carriers of that kind.

If two required carrier kinds were governed by a single undifferentiated regime,
then either the regime would not distinguish their identity behavior, or
it would distinguish them by consulting some further discriminator.
In the first case, one carrier kind could stand in for another, collapsing the reference inventory.
In the second case, the discriminator would be doing the identity work outside the regime structure.
That is a hidden regime, excluded by Section~\ref{sec:formal_setting}.
Therefore each required carrier kind contributes at least one identity regime.
\end{proof}

The inventory alone therefore gives a floor of six regimes.
This is not yet the main result.
The main result comes from the fact that three required carrier kinds
cannot be served by only one regime each.

% -----------------------------------------------------------------------------
\subsection{Witnessed Splits}
\label{subsec:witnessed_splits}
% -----------------------------------------------------------------------------

Three carrier kinds exhibit witnessed split pressure under the transformation basis.

First, plain referents split into locus-fixed and object-fixed identity.
By Proposition~\ref{se200.prop.PlainReferentUnderdetermination}, branch/fork separates these bases.
A fork may preserve the locus while breaking object identity.
Thus plain referents require both $\LOC$ and $\OBJ$.

Second, scopes split into extension-fixed and structure-fixed identity.
By Proposition~\ref{se200.prop.ScopeUnderdetermination}, aggregation/decomposition separates these bases.
A decomposition may preserve the covered cases while changing the internal scope structure.
Thus scopes require both $\SCOPEE$ and $\SCOPES$.

Third, rules split into content-fixed and structure-fixed identity.
By Proposition~\ref{se200.prop.RuleUnderdetermination}, refinement separates these bases.
A refinement may preserve what the rule requires while changing its textual or internal structure.
Thus rules require both $\RULEC$ and $\RULES$.

Each split adds one regime beyond the one-regime-per-carrier minimum.
The additional regimes are not additional reference kinds.
They are distinct identity bases for required carrier kinds.


% -----------------------------------------------------------------------------
\subsection{The Bound}
\label{subsec:the_bound}
% -----------------------------------------------------------------------------

\begin{theorem}[Nine-regime lower bound]
\label{se200.thm.NineRegimeLowerBound}
Under the assumptions of Section~\ref{sec:formal_setting},
a neutral accountability substrate operating under persistent disagreement
must realize at least nine identity regimes
for the carrier-basis pairs considered in this paper.
\end{theorem}

\begin{proof}
By Lemma~\ref{se200.lem.InventoryFloor},
the six required carrier kinds in Section~\ref{sec:inventory}
|require at least six identity regimes.

Three of those carrier kinds require an additional regime because they exhibit witnessed split pressure.
Plain referents require both $\LOC$ and $\OBJ$ by Proposition~\ref{se200.prop.PlainReferentUnderdetermination}.
Scopes require both $\SCOPEE$ and $\SCOPES$ by Proposition~\ref{se200.prop.ScopeUnderdetermination}.
Rules require both $\RULEC$ and $\RULES$ by Proposition~\ref{se200.prop.RuleUnderdetermination}.

Therefore the required lower-bound core contains:
\[
\begin{gathered}
\OBL,\quad \OCC,\quad \REC,\quad \LOC,\quad \OBJ,\\
\SCOPEE,\quad \SCOPES,\quad \RULEC,\quad \RULES .
\end{gathered}
\]
Equivalently,
\[
1 + 1 + 1 + 2 + 2 + 2 = 9 .
\]
Thus any neutral accountability substrate satisfying the stated assumptions
|must realize at least nine identity regimes for these carrier-basis pairs.
This proves the lower-bound claim.
\end{proof}

% -----------------------------------------------------------------------------
\subsection{Monotonicity}
\label{subsec:monotonicity}
% -----------------------------------------------------------------------------

\begin{corollary}[Monotonicity]
\label{se200.cor.Monotonicity}
Enlarging the transformation basis cannot lower the bound
established by Theorem~\ref{se200.thm.NineRegimeLowerBound}.
\end{corollary}

\begin{proof}
The nine-regime lower-bound core is witnessed by distinctions
already forced by the transformation basis $\Fset$.
Adding transformations adds classification coordinates.
It may leave the existing distinctions unchanged.
It may make an already-distinct pair distinct in additional ways.
It may also force further splits.
But it cannot erase a split already witnessed by $\Fset$.
Therefore any transformation basis containing $\Fset$ still requires
at least the nine core regimes derived here.
\end{proof}

The monotonicity claim is intentionally limited.
It establishes only that incompleteness of the transformation basis
cannot make the present lower bound disappear.
A richer basis may leave the bound unchanged or raise it, but cannot lower it.

% -----------------------------------------------------------------------------
\subsection{Limits}
\label{subsec:bound_not_claim}
% -----------------------------------------------------------------------------

The claim is a lower bound.
Under the stated neutral-substrate assumptions,
fewer than nine regimes cannot preserve stable reference for the carrier-basis pairs analyzed here.
The result does not say that these nine regimes exhaust the coherent identity regimes.
It does not say that the transformation basis is complete for every record system.
It does not say that a deployed system should expose these nine regimes and no others.

Additional admissible identity bases may force additional regimes.
For example, a record may be tracked by descriptive content,
by a particular file, by a signed original, by a cryptographic hash, or by a chain-of-custody token.
An occurrence may be tracked as a fine-grained realization or as a coarser event complex.
An obligation-bearer may be tracked as a person, role, or organizational unit.
Such refinements would raise the number of regimes required in a deployment.
They do not remove the nine-regime lower-bound core established here.
