% !TeX root = se200_identity_regimes.tex
\section{Conclusion}
\label{sec:conclusion}

This paper has addressed a structural question that arises when ontologies
are required to function as neutral substrates under persistent
disagreement.
Under explicit constraints of neutrality, stability under incompatible extension,
and the exclusion of hidden structural discriminations, we have
shown that a lower bound exists on the ontological structure
required to support accountability.
Ontologies with fewer than six distinct identity-and-persistence regimes
cannot satisfy the stated neutrality and stability requirements
under the assumptions fixed in Section~\ref{sec:assumptions}.
A construction with exactly these six regimes suffices without introducing hidden or excess structure.

The sufficiency construction shows that this boundary is achievable:
a substrate realizing exactly six identity-and-persistence regimes
satisfies all stated requirements without hidden or excess structure.
The result establishes not only that certain distinctions are unavoidable,
but that the minimal structure required to support
accountability under persistent disagreement is determinate.

The analysis shows that accountability under persistent disagreement
requires a cross-cutting stability constraint governing identity and
persistence across all represented entities (R1),
together with explicit representation of obligation-bearing entities, normative structures,
time-indexed occurrences, applicability contexts, and descriptive records (R2--R6),
each with distinct identity conditions.

The result is conditional rather than prescriptive.
It does not claim that the identified structure
is optimal for all applications,
nor does it assess the adequacy of
alternative modeling strategies for other objectives.
It establishes only that, under a particular and explicitly stated
optimization objective, certain distinctions are structurally unavoidable.

This work clarifies the design space for neutral substrates under persistent
disagreement.
Under the stated constraints, it shows that the
lower bound on explicit ontological structure is determined by structural
requirements rather than by modeling preference.
Future work may examine how alternative optimization objectives
yield different structural bounds,
or analyze how the six-regime construction behaves
when stability or neutrality assumptions are systematically weakened.
Such analyses may help clarify how different interoperability
objectives generate different minimal ontological structures.


\section*{Acknowledgements}

The author thanks the anonymous reviewers for their detailed and
technical engagement with the argument.
Their comments substantially strengthened the necessity proof
and clarified the scope of the central result.

Portions of this work were developed
using computer-assisted tools during manuscript preparation.
Generative language models were used to assist with editing, formatting,
and consistency checking.
All conceptual framing, formal development, results, interpretations, and conclusions
are the author's own.

The author reviewed all suggestions and
takes full responsibility for the content of this work.
