% !TeX root = ../se200_identity_regimes.tex

\subsection{Refinement of Normative Structures}
\label{subsec:nor_split}

This section shows that normative structures do not define a single
regime profile under the transformation basis, and must be refined.

%--------------------------------------------
\subsubsection{Initial NOR Profile}

\begin{definition}[Provisional NOR profile]
A provisional NOR profile is a regime profile whose identity carrier is
a normative structure, i.e., an enduring referent that grounds obligations
or permissions independently of any particular occurrence.
\end{definition}


Under this provisional formulation, identity is intended to track persistence
of a normative structure independently of representation or organization.


%--------------------------------------------
\subsubsection{Transformation Behavior}

We examine the classification of transformations under the provisional NOR profile.

\begin{itemize}
\item $\mathrm{RE}$: IGN
\item $\mathrm{AN}$: IGN
\item $\mathrm{RF}$: PRS or BRK
\item $\mathrm{AD}$: PRS or BRK
\item $\mathrm{RC}$: IGN
\item $\mathrm{RA}$: IGN
\item $\mathrm{SU}$: BRK
\item $\mathrm{BF}$: BRK
\item $\mathrm{PV}$: IGN
\item $\mathrm{SE}$: not applicable
      ($\mathrm{Appl}_P(\mathrm{SE}) = 0$;
      normative structures do not undergo state evolution).
\end{itemize}


The classification of refinement is indeterminate under the provisional NOR profile:
it is identity-preserving when normative content is unchanged,
but identity-breaking when structural organization changes.
This indeterminacy is the source of split pressure.


%--------------------------------------------
\subsubsection{Split Pressure}

\begin{proposition}[NOR split pressure]
\label{prop:nor_split_pressure}
The provisional NOR profile is under split pressure with respect to
the refinement transformation $\mathrm{RF}$.
\end{proposition}

\begin{proof}[Proof of Proposition~\ref{prop:nor_split_pressure}]
Let $x \in \mathcal{R}$ be a normative structure and let $\tau$ be a
refinement transformation producing a representation
$y := \tau(x)$ that preserves the same normative content.

If identity is determined by normative content,
then $\mathrm{Class}_P(\mathrm{RF}) = \mathrm{PRS}$ is required,
i.e., refinement must be identity-preserving.

If identity is determined by structural organization,
then $\mathrm{Class}_P(\mathrm{RF}) = \mathrm{BRK}$ is required,
i.e., refinement must be identity-breaking.

These requirements are contradictory.
No single value of $\mathrm{Class}_P(\mathrm{RF})$ is consistent with both cases.
Therefore the provisional NOR profile is under split pressure with respect to $\mathrm{RF}$.
\end{proof}

[Informally]
After refinement, the same normative content can remain
even when the structure does not.


%--------------------------------------------
\subsubsection{Separated Profiles}

We therefore define two distinct regime profiles.

\begin{definition}[NOR-C profile]
The content-preserving normative profile $\mathrm{NOR\mbox{-}C}$ is defined by:
\begin{itemize}
\item carrier: normative structure,
\item identity basis: persistence of the same normative content,
\item transformation classification:
\begin{itemize}
\item $\mathrm{RE}$: IGN
\item $\mathrm{AN}$: IGN
\item $\mathrm{RF}$: PRS
\item $\mathrm{AD}$: PRS
\item $\mathrm{RC}$: IGN
\item $\mathrm{RA}$: IGN
\item $\mathrm{SU}$: BRK
\item $\mathrm{BF}$: BRK
\item $\mathrm{PV}$: IGN
\item $\mathrm{SE}$: not applicable
      ($\mathrm{Appl}_P(\mathrm{SE}) = 0$;
      normative structures do not undergo state evolution).
\end{itemize}
\end{itemize}
\end{definition}

\begin{definition}[NOR-S profile]
The structural normative profile $\mathrm{NOR\mbox{-}S}$ is defined by:
\begin{itemize}
\item carrier: normative structure,
\item identity basis: persistence of the same structure,
\item transformation classification:
\begin{itemize}
\item $\mathrm{RE}$: IGN
\item $\mathrm{AN}$: IGN
\item $\mathrm{RF}$: BRK
\item $\mathrm{AD}$: BRK
\item $\mathrm{RC}$: IGN
\item $\mathrm{RA}$: IGN
\item $\mathrm{SU}$: BRK
\item $\mathrm{BF}$: BRK
\item $\mathrm{PV}$: IGN
\item $\mathrm{SE}$: not applicable
      ($\mathrm{Appl}_P(\mathrm{SE}) = 0$;
      normative structures do not undergo state evolution).
\end{itemize}
\end{itemize}
\end{definition}


Both $\mathrm{NOR\mbox{-}C}$ and $\mathrm{NOR\mbox{-}S}$ share the same carrier
(normative structure). The refinement is by identity basis rather than carrier type:
normative content versus structural organization.


%--------------------------------------------
\subsubsection{Non-Collapse}

\begin{proposition}[NOR-C and NOR-S do not collapse]
\label{prop:nor_noncollapse}
The profiles $\mathrm{NOR\mbox{-}C}$ and $\mathrm{NOR\mbox{-}S}$ induce
distinct equivalence relations on $\mathcal{R}$.
\end{proposition}

\begin{proof}[Proof of Proposition~\ref{prop:nor_noncollapse}]
Let $x \in \mathcal{R}$ be a normative structure and let $\tau$ be a
refinement transformation producing a representation
$y := \tau(x)$ with the same normative content.

Under $\mathrm{NOR\mbox{-}C}$:
\[
\mathrm{Class}_{\mathrm{NOR\mbox{-}C}}(\mathrm{RF}) = \mathrm{PRS},
\]
so $\tau$ is identity-preserving and $x \sim_{\mathrm{NOR\mbox{-}C}} y$.

Under $\mathrm{NOR\mbox{-}S}$:
\[
\mathrm{Class}_{\mathrm{NOR\mbox{-}S}}(\mathrm{RF}) = \mathrm{BRK},
\]
so $\tau$ is identity-breaking and $x \not\sim_{\mathrm{NOR\mbox{-}S}} y$.

Therefore $\sim_{\mathrm{NOR\mbox{-}C}} \neq \sim_{\mathrm{NOR\mbox{-}S}}$.
\end{proof}

[Informally]
After refinement, two representations can express the same normative content
even when their structure differs.


%--------------------------------------------
\subsubsection{Result}

\begin{theorem}[NOR refinement]
Normative structures must be refined into at least two distinct regime profiles:
\[
\mathrm{NOR\mbox{-}C}
\quad\text{and}\quad
\mathrm{NOR\mbox{-}S}.
\]
\end{theorem}

[Informally]
Two representations can encode the same normative content even when organized differently,
but cannot be the same structure once that organization changes.

